---
title: "105-alternative-fish"
output: html_document
date: "2025-09-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message=FALSE)
```

```{r}
library(tidyverse)
library(pcadapt)
library(viridis)
library(ggplotify)
library(pheatmap)
library(ggrepel)
library(ggpubr)
library(grid)
library(raster)
library(ozmaps)
library(ggrepel)
library(ggspatial)
```

Include, Daly, Roper, FitzroyQLD

```{r}
qld<-read_csv("meta/qld-fish.csv") %>% mutate(NewPhenotype=Phenotype) %>% filter(Group!="Mainstream")
qld
```


```{r}
mq<-read_csv("~/github/barra-lh/meta/meta44.csv") %>% mutate(Group="Burdekin") %>% bind_rows(qld) %>% filter(Coverage > 8)
mq$Group<-factor(mq$Group, levels=c("Daly River","Roper River","Mainstream","Burdekin","FitzroyQLD"))
ggplot(mq) +
  geom_histogram(aes(x=Coverage, fill=Group)) +
  scale_fill_viridis_d(option="magma") +
  theme_bw() +
  ylab("Count\n") +
  xlab("\nCoverage") 
```

```{r}
mq %>% group_by(Group) %>% summarise(Mean=mean(Coverage), Min=min(Coverage), Max=max(Coverage), Median=median(Coverage), Count=n())
```

Should probably downsample these...


```{r}
mq %>% dplyr::select(Path) %>% write_tsv("~/github/barra-lh/bamlists/69.bamlist", col_names = FALSE)
mq %>% dplyr::select(Pheno) %>% write_tsv("~/github/barra-lh/bamlists/69.pheno", col_names = FALSE)
mq %>% dplyr::select(specimen_id) %>% write_tsv("~/github/barra-lh/bamlists/69.names", col_names = FALSE)
write_csv(mq, file="meta/mq.csv")
meta<-mq
```



Call snps on all chroms

```{sh, eval=FALSE}
srun -t 5-24:00:00 -p high --mem=32G --nodes=1 --ntasks-per-node=1 --cpus-per-task=12 $HOME/angsd/angsd -nthreads 12 \
-minInd 66 -bam bamlists/69.bamlist -ref genome/GCF_001640805.2_TLL_Latcal_v3_genomic.fna \
-rf meta/lates-lgs.txt -out outputs/105/snps-wgs-05-glf  \
-minMaf 0.05 -minMapQ 20 -minQ 20 -GL 1 -doGLF 2 -doMajorMinor 1 -doMaf 1 -SNP_pval 1e-6 \
-doGeno 2 -doPost 1 -postCutoff 0.95 -doPlink 2  > outputs/105/snps-wgs-05-glf.out 2> outputs/105/snps-wgs-05-glf.err &
```

1135370

```{sh, eval=FALSE}
plink --tped snps-wgs-05-glf.tped --tfam snps-wgs-05-glf.tfam --out plink-binary-05 --recode --allow-extra-chr --noweb
plink --ped plink-binary-05.ped --map plink-binary-05.map --recode vcf --allow-extra-chr -out plink-05
bgzip plink-05.vcf 
tabix plink-05.vcf.gz

#renaming vcf bcftools/1.13
module load bcftools

bcftools reheader --samples bamlists/69.names -o outputs/105/renamed-05.vcf.gz outputs/105/plink-05.vcf.gz


bcftools +fill-tags outputs/105/renamed-05.vcf.gz  -- -t MAF,F_MISSING | bcftools view -i 'F_MISSING < 0.05' | bcftools view -i 'INFO/MAF > 0.05' |  bcftools +prune -m 0.20 -w 50kb --nsites-per-win 1 --nsites-per-win-mode rand  > outputs/105/pruned-05-rand.vcf

bcftools +fill-tags outputs/105/renamed-05.vcf.gz  -- -t MAF,F_MISSING | bcftools view -i 'F_MISSING < 0.05' | bcftools view -i 'INFO/MAF > 0.05' |  bcftools +prune -m 0.20 -w 50kb --nsites-per-win 1 --nsites-per-win-mode maxAF  > outputs/105/pruned-05-max.vcf

#maxAF or rand option
bcftools +fill-tags outputs/105/renamed-05.vcf.gz  -- -t MAF,F_MISSING | bcftools view -i 'F_MISSING < 0.05' | bcftools view -i 'INFO/MAF > 0.10' |  bcftools +prune -m 0.20 -w 50kb --nsites-per-win 1 --nsites-per-win-mode rand  > outputs/105/pruned-10.vcf


bgzip outputs/105/pruned-05-rand.vcf
tabix outputs/105/pruned-05-rand.vcf.gz

bgzip outputs/105/pruned-05-max.vcf
tabix outputs/105/pruned-05-max.vcf.gz

bgzip outputs/105/pruned-10.vcf
tabix outputs/105/pruned-10.vcf.gz
```




## Outlier clustering


```{r}
meta<-mq
outliers<-read_csv("outputs/104/outliers.csv")
ps<-read_csv("outputs/104/ps.csv")
ddf<-read_tsv("outputs/105/snps-wgs-05-glf.geno.gz", col_names = c("FALSE"))  %>% dplyr::select(-X72) %>% filter(`FALSE`=="NC_066856.1")
ddf[ddf==-1]<-NA
ddf2<-ddf %>% filter(X2 %in% ps$Position) 
m<-ddf2 %>% dplyr::select(-`FALSE`, -X2) %>% as.matrix()
annot<-data.frame(sample=as.character(colnames(m))) %>% column_to_rownames("sample")
annot$Phenotype<-meta$Phenotype
annot
annot_colors=list(Phenotype=c(SFF="red",SSS="blue",UNK="grey75"))

```
 color=colorRampPalette(c("#000004FF","#B63679FF","#FCFDBFFF"))))
 
```{r}

pheatmap(m, annotation_col = annot, cluster_rows = FALSE,  annotation_colors = annot_colors, color=colorRampPalette(viridis(3, option="cividis", direction=-1, alpha = 0.75))(3),
                   labels_col = meta$specimen_id, labels_row = ddf2$X2, fontsize_row = 14, fontsize_col = 6)
grid.text("foo", x=-20, rot=90, gp=gpar(fontsize=25))

```


```{r}
as.ggplot(
pheatmap(m, annotation_col = annot, cluster_rows = FALSE,  annotation_colors = annot_colors, color=colorRampPalette(viridis(3, option="cividis", direction=-1, alpha = 0.75))(3),
                   labels_col = meta$specimen_id, labels_row = ddf2$X2, fontsize_row = 14, fontsize_col = 6,
         angle_col = 315)) +
         
#  ggtitle("Heatmap of associated barra variants on Lca24\n") +
  labs(caption="\nIndividual") +
  theme(plot.title = element_text(hjust=0.5)) +
  theme(plot.caption = element_text(size=14, hjust=0.4, face="bold")) 

ggsave("outputs/105/top-associated-variants.jpeg", width=12, height=6)
ggsave("outputs/105/top-associated-variants.pdf", width=12, height=6)

```




23 SSS 
21 SFF

Misclassified SS:  
SFLc0683, SFLc0398

Misclassified SFF:
SFLc0054, SFLc0096, SFLc0714, 
SFLc0403, SFLc0356, SFLc0282

86% correct assignment


clustering_method = "ward.D"
```{r}
as.ggplot(
pheatmap(m, annotation_col = annot, cluster_rows = FALSE,  annotation_colors = annot_colors, color=colorRampPalette(viridis(3, option="cividis", direction=-1, alpha = 0.75))(3),
                   labels_col = meta$specimen_id, labels_row = ddf2$X2, fontsize_row = 14, fontsize_col = 6, clustering_method = "ward.D",
         angle_col = 315)) +
         
#  ggtitle("Heatmap of associated barra variants on Lca24\n") +
  labs(caption="\nIndividual") +
  theme(plot.title = element_text(hjust=0.5)) +
  theme(plot.caption = element_text(size=14, hjust=0.4, face="bold")) 

ggsave("outputs/105/top-associated-variants-ward.jpeg", width=12, height=6)
ggsave("outputs/105/top-associated-variants-ward.pdf", width=12, height=6)

```

```{r}
meta<-mq
ps<-read_csv("outputs/104/lca16.csv")
ddf<-read_tsv("outputs/105/snps-wgs-05-glf.geno.gz", col_names = c("FALSE"))  %>% dplyr::select(-X72) %>% filter(`FALSE`=="NC_066848.1")
ddf[ddf==-1]<-NA
ddf2<-ddf %>% filter(X2 %in% ps$Position) 
m<-ddf2 %>% dplyr::select(-`FALSE`, -X2) %>% as.matrix()
annot<-data.frame(sample=as.character(colnames(m))) %>% column_to_rownames("sample")
annot$Phenotype<-meta$Phenotype
annot
annot_colors=list(Phenotype=c(SFF="#F0978D",SSS="#63D7DE",UNK="grey50"))

```

```{r}
as.ggplot(pheatmap(m, annotation_col = annot, cluster_rows = FALSE,  annotation_colors = annot_colors,
                   labels_col = meta$specimen_id, labels_row = ddf2$X2, fontsize_row = 14, fontsize_col = 6)) +
  ggtitle("Heatmap of associated barra variants on Lca16\n") + 
  theme(plot.title = element_text(hjust=0.5))

ggsave("outputs/105/top-associated-variants-lca16.jpeg", width=12, height=6)
ggsave("outputs/105/top-associated-variants-lca16.pdf", width=12, height=6)

```

## PCA

```{r}
path_to_file <- "outputs/105/pruned-05-max.vcf.gz"

filename <- read.pcadapt(path_to_file, type = "vcf")
```

_1_ choose a K   

```{r}
x <- pcadapt(input = filename, K = 20) 
```

```{r}
var<-round(x$singular.values^2*100,2)
var
```

10% with max AF pc1
8.43 with rand snp pc1

```{r}
#plot(x, option = "screeplot")
panela<-plot(x, option = "screeplot", K = 10) + ggtitle("A") +
  theme_bw() +
  theme(plot.title = element_text(size=14, face="bold")) +
  theme(panel.grid = element_blank(), panel.background = element_blank()) +
  scale_x_continuous(breaks=seq(1,10,1), labels=seq(1,10,1)) +
  ylab("Proportion of Explained Variance\n") +
  xlab("\nPrincipal Component") +
  theme(axis.title=element_text(size=12, face="bold"))
panela

ggsave("outputs/105/scree-plot.jpeg", width=4, height=4)
```


```{r}
plot(x, option = "scores")
```


```{r}
meta<-mq
m<-meta
```

```{r}
pcadata<-x$scores[,1:5] %>% as_tibble()
pcadata<-pcadata %>% bind_cols(m)

```

```{r}
p1<-ggplot(pcadata) +
  geom_point(aes(x=V1, y=V2, fill=NewPhenotype, shape=Group), alpha=0.75, cex=2) +
  theme_bw() +
 # theme(legend.position = "top") +
  theme(panel.grid = element_blank()) +
  xlab("PC 1") +
  ylab("PC 2") +
#  scale_fill_viridis_d(option="H") +
  scale_fill_manual(values=c("red","blue","grey")) +
  scale_shape_manual(values=c(24,22,21,23)) +
  xlab(paste0("PC1 ",var[1],"%")) +
  ylab(paste0("PC2 ",var[2],"%")) +
  theme(axis.title=element_text(size=12, face="bold"))


p1

```


```{r}
p2<-ggplot(pcadata) +
  geom_point(aes(x=V1, y=V3, fill=NewPhenotype, shape=Group), alpha=0.75, cex=2) +
  theme_bw() +
  #theme(legend.position = "none") +
  theme(panel.grid = element_blank()) +
#  scale_fill_viridis_d(option="H") +
  scale_fill_manual("Migratory Phenotype",values=c("red","blue","grey"), labels=c("MIG","BWR","UNK"), ) +
  scale_shape_manual("Sampling Location",values=c(24,22,21,23), labels=c("Daly River","Roper River","Burdekin Region","Fitzroy River")) +
  xlab(paste0("PC1 ",var[1],"%")) +
  ylab(paste0("PC3 ",var[3],"%")) + 
  guides(fill=guide_legend(override.aes=list(shape=21))) +
    theme(axis.title=element_text(size=12, face="bold"))


p2
```

```{r}
ggplot(pcadata) +
  geom_point(aes(x=V1, y=V4, fill=NewPhenotype, shape=Group), alpha=0.75, cex=2) +
  theme_bw() +
  #theme(legend.position = "none") +
  theme(panel.grid = element_blank()) +
#  scale_fill_viridis_d(option="H") +
  scale_fill_manual("Migratory Phenotype",values=c("red","blue","grey"), labels=c("MIG","BWR","UNK"), ) +
  scale_shape_manual("Sampling Location",values=c(24,22,21,23), labels=c("Daly River","Roper River","Burdekin Region","Fitzroy River")) +
  xlab(paste0("PC1 ",var[1],"%")) +
  ylab(paste0("PC4 ",var[4],"%")) + 
  guides(fill=guide_legend(override.aes=list(shape=21))) +
    theme(axis.title=element_text(size=12, face="bold"))
```

```{r}
ggarrange(p1+theme(legend.position = "none"),p2, widths=c(1,1.3))

ggsave("outputs/105/new-australian-data-pcs.jpeg", width=10, height=4)
```

## Let's pull out outliers on PC1 and see what they are!! Are the Life-history variants part of it?



```{r}
x <- pcadapt(filename, K = 4)
summary(x)
```

```{r}
plot(x , option = "manhattan")
```

```{r}
plot(x, option = "qqplot")
```

```{r}
hist(x$pvalues, xlab = "p-values", main = NULL, breaks = 50, col = "orange")
```

```{r}
plot(x, option = "stat.distribution")
```


```{R}
par(mfrow = c(2, 2))
for (i in 1:4)
  plot(x$loadings[, i], pch = 19, cex = .3, ylab = paste0("Loadings PC", i))
```    
```{r}
plot(x)
```

```{r}
padj <- p.adjust(x$pvalues,method="bonferroni")
alpha <- 0.05
outliers <- which(padj < alpha)
length(outliers)

snp_pc <- get.pc(x, outliers)
snp_pc %>% group_by(PC) %>% summarize(Count=n())

```

```{r}
padj <- p.adjust(x$pvalues,method="BH")
alpha <- 0.05
outliers <- which(padj < alpha)
length(outliers)

snp_pc <- get.pc(x, outliers)
snp_pc %>% group_by(PC) %>% summarize(Count=n())

```


get snps `gunzip -c outputs/105/pruned-05-max.vcf.gz| grep -v "^#" | cut -f1-3 | awk '{print $0"\t"NR}' > outputs/105/pruned-05-max-snp-numbers.tsv`


```{r}
snp_pc1<-snp_pc %>% filter(PC==1)
labels<-read_tsv("meta/lates-lgs-named.txt") %>% rename(Name=Chrom, Chrom=Chromosome)
labels
```


```{r}
snpnumbers<-read_tsv("outputs/105/pruned-05-max-snp-numbers.tsv", col_names = c("Chrom","Pos","Marker","SNP"))
snps<-left_join(as_tibble(snp_pc1), snpnumbers) 

snps %>% group_by(PC, Chrom) %>% summarize(Count=n()) %>% arrange(-Count)

snps %>% left_join(labels) 
```
# Map of Australia, with states and sampled populations

## Get meta

```{r}
sites<-read_tsv("meta/sites.tsv")
states<-read_tsv("meta/state-data.txt")
states
```
```{r}
domain <- c(105, 158, -46, -5)
```

```{r}
oz <- ozmap_data("states")
locs1<-ggplot(oz) +
  geom_sf() +
  geom_point(data=sites, aes(x=long, y=lat), shape=sites$shape, fill=sites$color, size=5) +
  geom_text_repel(data=sites, aes(x=long, y=lat, label=Locality)) +
  annotation_scale() +
     annotation_north_arrow(location = "bl", which_north = "true", 
        pad_x = unit(0.2, "in"), pad_y = unit(0.3, "in"), style=north_arrow_nautical())

locs1
```
  geom_text_repel(data=aus,aes(x=Longitude, y=Latitude, label=Host), max.overlaps = Inf) +
  geom_text(data=states, aes(x=Longitude, y=Latitude, label=State)) +
  coord_sf(xlim = c(domain[1]+4,domain[2]-2.5), ylim=c(domain[3]+2, domain[4]-5)) +
 +
  scale_fill_manual(values=colors[1:4]) +
  theme(legend.position = "none")

## Get raster
```{r, eval=FALSE}
nat.earth<-stack("~/github/fish-lake/data/earth/NE2_HR_LC_SR_W_DR/NE2_HR_LC_SR_W_DR.tif")

nat.crop <- crop(nat.earth, y=extent(domain))

rast.table <- data.frame(xyFromCell(nat.crop, 1:ncell(nat.crop)),
                         getValues(nat.crop/255))


rast.table$rgb <- with(rast.table, rgb(NE2_HR_LC_SR_W_DR_1,
                                       NE2_HR_LC_SR_W_DR_2,
                                       NE2_HR_LC_SR_W_DR_3,
                                       1))

save(rast.table, file="data/rast.rda")

# here eric implements something and I copy it
tidy_subset <- function(x, longlat) {
  x@data$id <- rownames(x@data)
  x.f <- broom::tidy(x) %>%
    dplyr::left_join(., x@data, by = "id") %>%
    dplyr::tbl_df() %>%
    filter(long > longlat[1],
           long < longlat[2],
           lat > longlat[3],
           lat < longlat[4])
}

```

```{r}
load("data/rast.rda")
```


```{r}
locs<-ggplot() +
  geom_raster(data = rast.table, mapping = aes(x = x, y = y), fill = rast.table$rgb, interpolate = TRUE)  +
   geom_sf(data=oz, fill=NA) +
   geom_point(data=sites, aes(x=long, y=lat), shape=sites$shape, fill=sites$color, size=4) +
   geom_text_repel(data=sites, aes(x=long, y=lat, label=Locality)) +
   coord_sf(xlim = c(domain[1]+5,domain[2]-2.5), ylim=c(domain[3]+2, domain[4]-5))  +
   annotation_scale() +
     annotation_north_arrow(location = "bl", which_north = "true", 
        pad_x = unit(0.2, "in"), pad_y = unit(0.3, "in"), style=north_arrow_nautical()) +
  ylab("Latitude") +
  xlab("Longitude") +
  theme_bw() +
  theme(axis.title = element_text(size=14, face="bold"))
locs

ggsave("outputs/105/map.jpeg", width=4, height=4)
```  
  geom_jitter(data=aus, aes(x=Longitude, y=Latitude, fill=Location), pch=25, size=3) +
  geom_label_repel(data=aus,aes(x=Longitude, y=Latitude, label=tip_name), max.overlaps = Inf, size=2, alpha=0.8) +
  geom_text(data=states, aes(x=Longitude, y=Latitude, label=State)) +

  theme(panel.grid=element_blank())  +
    scale_fill_manual(values=colors[1:4]) +
  theme(legend.position = "none") +
  annotation_scale() +
     annotation_north_arrow(location = "bl", which_north = "true", 
        pad_x = unit(0.2, "in"), pad_y = unit(0.3, "in"), style=north_arrow_nautical())
ggsave("outputs/300/plasmo-map-01312025.pdf", width=8, height=6)
ggsave("outputs/300/plasmo-map-01312025.jpeg", width=8, height=6)
```{r}
blank <- grid.rect(gp=gpar(col="white"))
```
```{r}
ab<-ggarrange(locs + ggtitle("A") + theme(plot.title = element_text(size=18, face="bold")),
            ggarrange(ggplotGrob(ggarrange(blank, panela + ggtitle("B")+ theme(plot.title = element_text(size=18, face="bold")), blank, ncol=1, heights=c(.5,2,.5))),blank, widths=c(2,1)), widths=c(1,1))

ab
```




```{r}
cd<-ggarrange(p1 + theme(legend.position = "none") + ggtitle("C") + theme(plot.title = element_text(size=18, face="bold")),
              p2 + ggtitle("D") + theme(plot.title = element_text(size=18, face="bold")), widths=c(1,1.4))
cd
```


```{r}
ggarrange(ab, cd, ncol=1)

ggsave("outputs/105/combined-figure.pdf", width=10, height=8)
ggsave("outputs/105/combined-figure.jpeg", width=10, height=8)

```