---
title: "111-kinship"
output: html_document
date: "2026-03-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```


```{r}
library(tidyverse)
library(viridis)
```

Testing for kin in the Burdekin barra
```{sh, eval=FALSE}
# https://github.com/ANGSD/NgsRelate
### First we generate a file with allele frequencies (angsdput.mafs.gz) and a file with genotype likelihoods (angsdput.glf.gz).

srun -p bmh -t 16:00:00 --mem=32G --nodes=1 --cpus-per-task=6  $HOME/angsd/angsd -P 6  \
-b $HOME/barra-lh/bamlists/44.bamlist -rf meta/lates-lgs.txt \
-ref  genome/GCF_001640805.2_TLL_Latcal_v3_genomic.fna \
-minInd 42 -minMapQ 20 -minQ 20 \
-GL 2 -doMajorMinor 1 -doMaf 1 -snp_pval 1e-6 -minMaf 0.05 -doGlf 3 \
-out outputs/111/kin > outputs/111/kin.out 2> outputs/111/kin.err &

#	-> Number of sites retained after filtering: 1609596 
#	[ALL done] cpu-time used =  43680.16 sec
#	[ALL done] walltime used =  24467.00 sec
### Then we extract the frequency column from the allele frequency file and remove the header (to make it in the format NgsRelate needs)

zcat outputs/111/kin.mafs.gz | cut -f6 |sed 1d > outputs/111/freq

### run NgsRelate
module load ngsrelate/2.0     
srun -p bmh -t 12:00:00 --mem=32G --nodes=1 --cpus-per-task=2 ngsRelate -g outputs/111/kin.glf.gz -n 44 -f outputs/111/freq  -O outputs/111/kin-res.tsv

#        [ALL done] cpu-time used =  4531.40 sec (filereading took: 7.68 sec)
#        [ALL done] walltime used =  2299.00 sec (filereading took: 8.00 sec)

```

```{r}
m44<-read_csv("meta/meta44.csv") %>% mutate(SampleID=1:n())
m44$SampleID<-m44$SampleID-1
m44
mkin<-m44
```

```{r}
rel<-read_tsv(file = "outputs/111/kin-res.tsv") %>% left_join(mkin, by=c("a"="SampleID")) %>% 
  rename(SampleNameA=specimen_id) %>%
  left_join(mkin, by=c("b"="SampleID")) %>% 
  rename(SampleNameB=specimen_id) %>% relocate(SampleNameA, SampleNameB)

rel %>% arrange(-theta) %>% select(SampleNameA, SampleNameB, rab, theta) %>% 
  write_csv("outputs/111/kin.csv")

rel %>% arrange(-theta) %>% select(SampleNameA, SampleNameB, rab, theta) %>% 
  filter(rab>0.15)
```


```{r}
cols<-viridis(6, alpha=0.75)
```

```{r}
ggplot(rel)+geom_histogram(aes(x=rab), bins=50, fill="grey", color="grey") +
  geom_vline(xintercept =.50, col=cols[1]) + #Parent offspring/full sibs
  geom_vline(xintercept =.25, col=cols[2]) + #2nd degree half sib, grandparent, aunt/uncle, double first cousins
  geom_vline(xintercept =.125, col=cols[3]) + # 3rd cousins
  geom_vline(xintercept = .0625, col=cols[4]) + 
  geom_vline(xintercept = .03125, col=cols[5]) +
  geom_vline(xintercept = 0.0, col=cols[6]) +
  scale_x_continuous(breaks=c(0,.03125, .0625, .125, .25, .5),
                     labels=c("0 / Unrelated","5th Deg.","4th Deg.","3rd Deg.", "2nd Deg.", "1st Deg.")) +
  theme_bw() +
  theme(axis.text.x = element_text(hjust=1, angle=45)) +
  theme(panel.grid=element_blank()) +
  ylab("Count") +
  xlab("Pairwise Relatedness") 
```

```{r}
ggplot(rel)+geom_histogram(aes(x=theta), bins=50, fill="grey", color="grey") +
  geom_vline(xintercept =.25, col=cols[1]) + #Parent offspring/full sibs
  geom_vline(xintercept =.125, col=cols[2]) + #2nd degree
  geom_vline(xintercept =.0625, col=cols[3]) + # 3rd
  geom_vline(xintercept = .0313, col=cols[4]) +
  geom_vline(xintercept = .0156, col=cols[5]) +
  geom_vline(xintercept = 0.0, col=cols[6]) +
  scale_x_continuous(breaks=c(0,0.0156, 0.0313, 0.0625, 0.125, 0.25),
                     labels=c("0 / Unrelated","5th Deg.","4th Deg.","3rd Deg.", "2nd Deg.", "1st Deg.")) +
  theme_bw() +
  theme(axis.text.x = element_text(hjust=1, angle=45)) +
  theme(panel.grid=element_blank()) +
  ylab("Count") +
  xlab("Kinship Coefficient")
```


```{r}
rel %>% filter(theta > 0.125) %>% select(SampleNameA, SampleNameB, theta)
```


Pretty similar results with close actually
SampleNameA SFLc0650 SFLc0428 SFLc0427 SFLc0403 SFLc0356 SFLc0398 SFLc0435 SFLc0228
pivot_wider(close3, id_cols = SampleNameA, names_from = SampleNameB, values_from = theta)
```{r}
close<-rel %>% select(SampleNameA, SampleNameB, theta) %>% arrange(-theta)
close2<-close %>% rename(b=SampleNameB, a=SampleNameA)
close2<-close2 %>% rename(SampleNameA=b, SampleNameB=a)
close3<-bind_rows(close,close2)
```


```{r}
ggplot(close3) +
  geom_tile(aes(x=SampleNameA, y=SampleNameB,fill=theta)) +
  scale_fill_viridis_c()
```

```{r}
dr<-pivot_wider(close3 %>% arrange(SampleNameA), id_cols = SampleNameA, names_from = SampleNameB, values_from = theta)
rownames<-dr$SampleNameA     
dr<-dr %>% select(-SampleNameA)
#dr[is.na(dr)]<-1

```

```{r}
annotation<-data.frame(sample=as.character(colnames(dr))) %>% column_to_rownames("sample")
annotation$Phenotype<-m44$NewPhenotype
annotationcolors=list(Phenotype=c(SFF="red",SSS="blue",UNK="grey75"))

ggplotify::as.ggplot(pheatmap(as.matrix(dr), labels_row = rownames, cluster_rows = TRUE, fontsize_row = 6, fontsize_col = 6,
                              annotation_col = annotation, annotation_colors =  annotationcolors ))

ggsave("outputs/111/kinship-matrix.jpeg", width=11, height=8)
```
```{r}
rel %>% filter(theta > 0.0625) %>% select(SampleNameA, SampleNameB, theta) %>% arrange(-theta)
```
SFLc0228, SFLc0427, SFLc0428, SFLc0356, SFLc0403, SFLc0398 are a potential family
2 SSS 4 SFF fish 

SFLc0435	SFLc0650
are another
```{r}
a<-rel %>% filter(theta > 0.0625) %>% select(SampleNameA, SampleNameB, theta) %>% arrange(-theta) %>% select(SampleNameA)
b<-rel %>% filter(theta > 0.0625) %>% select(SampleNameA, SampleNameB, theta) %>% arrange(-theta) %>% select(SampleNameB)
unique(c(a$SampleNameA, b$SampleNameB))
```

## Let's make some clones and see if that fixes this
SFLc0096 614649
SFLc0312 614660
SFLc0619 614683
in data/kin
```{sh, eval=FALSE}
cp ../bams/616648.sort.flt.bam Dup1.bam
cp ../bams/616650.sort.flt.bam Dup2.bam
cp ../bams/616652.sort.flt.bam Dup3.bam
for f in *.bam; do samtools index $f; done;

samtools merge -o Comb12.bam Dup1.bam Dup2.bam; samtools merge -o Comb13.bam Dup1.bam Dup3.bam; samtools merge -o Comb23.bam Dup2.bam Dup3.bam; 

for f in Comb*.bam; do samtools index $f; done;

```

Testing for kin in the Burdekin barra including some intentional dups.
```{sh, eval=FALSE}
# https://github.com/ANGSD/NgsRelate
### First we generate a file with allele frequencies (angsdput.mafs.gz) and a file with genotype likelihoods (angsdput.glf.gz).

srun -p bmh -t 16:00:00 --mem=32G --nodes=1 --cpus-per-task=6  $HOME/angsd/angsd -P 6  \
-b $HOME/barra-lh/bamlists/kin.bamlist -rf meta/lates-lgs.txt \
-ref  genome/GCF_001640805.2_TLL_Latcal_v3_genomic.fna \
-minInd 48 -minMapQ 20 -minQ 20 \
-GL 2 -doMajorMinor 1 -doMaf 1 -snp_pval 1e-6 -minMaf 0.05 -doGlf 3 \
-out outputs/111/kin-dup > outputs/111/kin-dup.out 2> outputs/111/kin-dup.err &

#        -> Number of sites retained after filtering: 1657786
 #       [ALL done] cpu-time used =  89829.79 sec
  #      [ALL done] walltime used =  48698.00 sec

zcat outputs/111/kin-dup.mafs.gz | cut -f6 |sed 1d > outputs/111/freq-dup

### run NgsRelate
module load ngsrelate/2.0     
srun -p bmh -t 12:00:00 --mem=32G --nodes=1 --cpus-per-task=2 ngsRelate -g outputs/111/kin-dup.glf.gz -n 50 -f outputs/111/freq-dup  -O outputs/111/kin-dup-res.tsv



```



```{r}
mkin<-read_csv("meta/kin.csv") %>% mutate(SampleID=1:n())
mkin$SampleID<-mkin$SampleID-1
mkin
```

```{r}
rel<-read_tsv(file = "outputs/111/kin-dup-res.tsv") %>% left_join(mkin, by=c("a"="SampleID")) %>% 
  rename(SampleNameA=specimen_id) %>%
  left_join(mkin, by=c("b"="SampleID")) %>% 
  rename(SampleNameB=specimen_id) %>% relocate(SampleNameA, SampleNameB)

rel %>% arrange(-theta) %>% select(SampleNameA, SampleNameB, rab, theta) %>% 
  write_csv("outputs/111/kin-dp.csv")

rel %>% arrange(-theta) %>% select(SampleNameA, SampleNameB, rab, theta, Fa, Fb) %>% 
  filter(rab>0.5)
```

```{r}
rel2 <- rel %>% filter(! SampleNameA %in% c("Dup1","Dup2","Dup3","Comb12","Comb13","Comb23"))
rel2 <- rel2 %>% filter(! SampleNameB %in% c("Dup1","Dup2","Dup3","Comb12","Comb13","Comb23"))

```

```{r}
cols<-viridis(6, alpha=0.75)
```

```{r}
ggplot(rel2)+geom_histogram(aes(x=rab), bins=50, fill="grey", color="grey") +
  geom_vline(xintercept =.50, col=cols[1]) + #Parent offspring/full sibs
  geom_vline(xintercept =.25, col=cols[2]) + #2nd degree half sib, grandparent, aunt/uncle, double first cousins
  geom_vline(xintercept =.125, col=cols[3]) + # 3rd cousins
  geom_vline(xintercept = .0625, col=cols[4]) + 
  geom_vline(xintercept = .03125, col=cols[5]) +
  geom_vline(xintercept = 0.0, col=cols[6]) +
  scale_x_continuous(breaks=c(0,.03125, .0625, .125, .25, .5),
                     labels=c("0 / Unrelated","5th Deg.","4th Deg.","3rd Deg.", "2nd Deg.", "1st Deg.")) +
  theme_bw() +
  theme(axis.text.x = element_text(hjust=1, angle=45)) +
  theme(panel.grid=element_blank()) +
  ylab("Count") +
  xlab("Pairwise Relatedness") 
```

```{r}
rel %>% filter(theta > 0.125) %>% select(SampleNameA, SampleNameB, theta)

```


```{r}
ggplot(rel2)+geom_histogram(aes(x=theta), bins=50, fill="grey", color="grey") +
  geom_vline(xintercept =.25, col=cols[1]) + #Parent offspring/full sibs
  geom_vline(xintercept =.125, col=cols[2]) + #2nd degree
  geom_vline(xintercept =.0625, col=cols[3]) + # 3rd
  geom_vline(xintercept = .0313, col=cols[4]) +
  geom_vline(xintercept = .0156, col=cols[5]) +
  geom_vline(xintercept = 0.0, col=cols[6]) +
  scale_x_continuous(breaks=c(0,0.0156, 0.0313, 0.0625, 0.125, 0.25),
                     labels=c("0 / Unrelated","5th Deg.","4th Deg.","3rd Deg.", "2nd Deg.", "1st Deg."), limits = c(0,.5)) +
  theme_bw() +
  theme(axis.text.x = element_text(hjust=1, angle=45)) +
  theme(panel.grid=element_blank()) +
  ylab("Count\n") +
  xlab("\nKinship Coefficient") 

ggsave("outputs/111/kinship-coefficient.jpeg")
```


```{r}
rel2 %>% filter(theta > .125) %>% select(SampleNameA, SampleNameB, theta, Fa, Fb)
```

SFLc0427	SFLc0428	0.141781	0.000000
SFLc0435	SFLc0650	0.196559  0e+00

```{r}
a<-rel %>% select(SampleNameA, Fa) %>% rename(Sample=SampleNameA, F=Fa)
b<-rel %>% select(SampleNameB, Fb) %>% rename(Sample=SampleNameB, F=Fb)

c<-bind_rows(a,b) %>% unique()
c
```
