---
title: "112-gwas-redo"
output: html_document
date: "2026-03-23"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

Following 

https://cloufield.github.io/GWASTutorial/04_Data_QC/#table-of-contents

and

https://www.staff.ncl.ac.uk/heather.cordell/FASTLMM-2025.html


We have a bunch of calls in plink and vcf format.
```{sh, eval=FALSE}
srun -t 4-24:00:00 -p high --mem=32G --nodes=1 --ntasks-per-node=1 --cpus-per-task=12 $HOME/angsd/angsd -nthreads 12 \
-minInd 42 -bam bamlists/44.bamlist -ref genome/GCF_001640805.2_TLL_Latcal_v3_genomic.fna \
-rf meta/lates-lgs.txt -out outputs/104/snps-wgs-05-glf  \
-minMaf 0.05 -minMapQ 20 -minQ 20 -GL 1 -doGLF 2 -doMajorMinor 1 -doMaf 1 -SNP_pval 1e-6 \
-doGeno 2 -doPost 1 -postCutoff 0.95 -doPlink 2  > outputs/104/snps-wgs-05-glf.out 2> outputs/104/snps-wgs-05-glf.err &
```


Moving snps-wgs-05-glf.t* locally
Moving plink-binary* locally
 plink-binary-05.ped
filtering heavily

```{sh, eval=FALSE}
plink \
    --ped plink-binary-05.ped --map plink-binary-05.map \
    --maf 0.05 \
    --geno 0.05 \
    --mind 0.05 \
    --hwe 1e-6 \
    --indep-pairwise 50kb 10 0.2 \
    --out plink-pruned \
    --allow-extra-chr --noweb
    
```

0 variants removed due to missing genotype data (--geno).
--hwe: 46441 variants removed due to Hardy-Weinberg exact test.
14585 variants removed due to minor allele threshold(s)
(--maf/--max-maf/--mac/--max-mac).
969724 variants and 44 people pass filters and QC.
Pruning complete.  882827 of 969724 variants removed.


F coefficient

```{sh, eval=FALSE}
plink \
    --ped plink-binary-05.ped --map plink-binary-05.map \
    --extract plink-pruned.prune.in \
    --het \
    --out plink_results \
    --allow-extra-chr --noweb

```

30838 variants scanned, 

awk 'NR>1 && $6>0.1 || $6<-0.1 {print $1,$2}' plink_results.het > high_het.sample

```{r}
f<-read_table("outputs/112/plink_results.het")

ggplot(f) +
  geom_histogram(aes(x=F))
```


outliers ?

```{r}
f %>% mutate(Mean=mean(F)) %>% mutate(StdDev=sd(F)) %>% mutate(Max=Mean+(3*StdDev)) %>%
  mutate(Min=Mean - (3*StdDev)) %>% filter(F < Min)

f %>% mutate(Mean=mean(F)) %>% mutate(StdDev=sd(F)) %>% mutate(Max=Mean+(3*StdDev)) %>%
  mutate(Min=Mean - (3*StdDev)) %>% filter(F > Max)
```

/home/maccamp/barra-lh/data/bams/616673.sort.flt.bam
616673 SFLc0427


awk 'NR>1 && $6>0.228 || $6<-0.567 {print $1,$2}' plink_results.het > high_het.sample


IBD
```{sh, eval=FALSE}
plink \
    --ped plink-binary-05.ped --map plink-binary-05.map \
    --extract plink-pruned.prune.in    \
    --genome \
    --out plink_results  \
    --allow-extra-chr --noweb
    
```

PI_HAT

```{r}
pi<-read_table("outputs/112/plink_results.genome")
pi$PI_HAT
```
Now, to create a clean data set

```{sh, eval=FALSE}
plink \
    --ped plink-binary-05.ped --map plink-binary-05.map \
    --maf 0.2 \
    --geno 0.05 \
    --mind 0.05 \
    --hwe 1e-6 \
    --remove high_het.sample \
    --keep-allele-order \
    --make-bed \
    --out sample_data.clean \
    --allow-extra-chr --noweb
```

466267 variants and 43 people pass filters and QC.
