---
title: "113-fst"
output: html_document
date: "2026-03-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message = FALSE)
```

We can also look at Fst windows

SFLc0427	SFLc0428	0.139075		
SFLc0435	SFLc0650	0.188466	

Are related individuals

Keeping SFLc0428 and SFLc0435 because they are higher coverage

```{r}
m44<-read_csv("meta/meta44.csv")
m42<-m44 %>% filter(!specimen_id %in% c("SFLc0427","SFLc0650"))
mig <- m42 %>% filter(NewPhenotype =="SFF")
res <- m42 %>% filter(NewPhenotype =="SSS")
m42 %>% select(specimen_id) %>% write_tsv("bamlists/42.names", col_names = FALSE)
write_csv(m42, "meta/m42.csv")


```

sample sizes are 21 & 21 with m42


```{r}
mig %>% select(Path) %>% write_tsv("bamlists/mig.bamlist", col_names = FALSE)
res %>% select(Path) %>% write_tsv("bamlists/res.bamlist", col_names = FALSE)
```

 bash 113-calc-sfs.sh 
 
 actually hiad this written up previously, using 202.1-calc-sfs.sh
 bash 202.1-calc-sfs.sh
$ scontrol requeue <job_id>
$ scontrol release <job_id>
$ scancel <job_id>


# With SNPR
We have some called snps
bcftools reheader --samples bamlists/44.names -o outputs/104/renamed-05.vcf.gz outputs/104/plink-05.vcf.gz

bcftools +fill-tags outputs/104/renamed-05.vcf.gz  -- -t MAF,F_MISSING | bcftools view -i 'F_MISSING < 0.05' | bcftools view -i 'INFO/MAF > 0.05' |  bcftools +prune -m 0.20 -w 50kb --nsites-per-win 1 --nsites-per-win-mode rand  > outputs/104/pruned-05.vcf
```{sh}
bcftools view outputs/104/renamed-05.vcf.gz -S bamlists/42.names | bcftools +fill-tags -- -t MAF,F_MISSING | bcftools view -i 'F_MISSING < 0.05' | bcftools view -i 'INFO/MAF > 0.05' > outputs/113/filt.vcf
bgzip outputs/113/filt.vcf
tabix outputs/113/filt.vcf.gz
bcftools view -Ov -r NC_066833.1,NC_066856.1 outputs/113/filt.vcf.gz > outputs/113/twochroms.vcf
```

996731 variants at maf05, the complete thing is too big for my current SNPr setup
Error: C stack usage 21536065 is too close to the limit
```{r, eval=FALSE}
#snps<-import.snpR.data("outputs/113/filt.vcf.gz", sample.meta = dplyr::select(m42,NewPhenotype,NewPheno,specimen_id) %>% rename(FishID=specimen_id))

snps<-import.snpR.data("outputs/113/twochroms.vcf", sample.meta = dplyr::select(m42,NewPhenotype,NewPheno,specimen_id) %>% rename(FishID=specimen_id))

snps@snp.meta$CHROM<-gsub("\\.1","",snps@snp.meta$CHROM)
p<-plot_clusters(snps, "NewPhenotype")
p$plots$pca

x<-calc_association(snps, response = "NewPheno", method = "chisq")
y<-get.snpR.stats(x, stats="association")


plot_manhattan(x, "chi_p_NewPheno", chr = "CHROM",
               log.p = TRUE)$plot

ggsave("outputs/104/snp-r-manhattan.jpeg")
```

```{r, eval=FALSE}
y$single %>% mutate(Position=1:n()) %>% group_by(CHROM) %>%  mutate(BHfdr = p.adjust(chi_p_NewPheno, method = "BH")) %>% arrange(BHfdr) %>% write_csv("outputs/104/snpr-gwas-results.csv")
```


sliding window fst

```{r}
fst<-calc_pairwise_fst(snps, facets="CHROM.NewPheno", method="wc")
fsts<-get.snpR.stats(fst, facets = "CHROM.NewPheno", stats = "fst")
```

```{r}
ggplot(fsts$pairwise) +
  geom_point(aes(x=position, y=fst))  +
  facet_wrap(.~CHROM)
```



```{r}
smooth<-calc_smoothed_averages(fst, facets = "CHROM.NewPheno", stats.type="pairwise", sigma=50, step=20)
smoothf<-get.snpR.stats(smooth, "CHROM.NewPheno", "pairwise.window") 
```


```{r}
ggplot(smoothf) +
  geom_point(aes(x=position, y=fst)) +
  facet_wrap(.~snp.subfacet)
```


```{r}
ggplot(smoothf) +
  geom_boxplot(aes(y=fst))
```