---
title: "115-outflank-more-stringent"
output: html_document
date: "2026-03-31"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Originally I kept SFLc0428 and SFLc0435 because they are higher coverage

SFLc0228, SFLc0427, SFLc0428, SFLc0356, SFLc0403 SFLc0398 are a potential family
2 SSS 4 SFF fish 

SFLc0435	SFLc0650
are another
1 SSS (0650) 1 SFF fish (0435)

```{r}
m44<-read_csv("meta/meta44.csv")
m44 %>% filter(specimen_id %in% c("SFLc0228","SFLc0427","SFLc0428","SFLc0356","SFLc0403",
                                  "SFLc0398")) %>% relocate(NewPhenotype)

m44 %>% filter(specimen_id %in% c("SFLc0435","SFLc0650")) %>% relocate(NewPhenotype)

```

Now I can filter to keep sampleSizes,

if we drop out kin, we end up 17 SFF and 20 SS (newphenotype), so we need to keep 2 SFF fish
Keeping SFLc0435 and SFLc0356 based on highest coverage
```{r}
#m44 %>% filter(!specimen_id %in% c("SFLc0228", "SFLc0427", "SFLc0428", "SFLc0356", "SFLc0403","SFLc0435","SFLc0650")) 
m38<-m44 %>% filter(!specimen_id %in% c("SFLc0228", "SFLc0427", "SFLc0428", "SFLc0403","SFLc0650","SFLc0398")) 

mig <- m38 %>% filter(NewPhenotype =="SFF") #19
res <- m38 %>% filter(NewPhenotype =="SSS") #19

write_csv(m38,"meta/m38.csv")
m38 %>% select(specimen_id) %>% write_tsv("bamlists/38.names", col_names = FALSE)
m38 %>% select(Path) %>% write_tsv("bamlists/38.bamlist", col_names=FALSE)
m38 %>% select(NewPheno) %>% write_tsv("bamlists/38.pheno",col_names = FALSE)
```


# Now to OutFlank

```{r}
chroms<-read_tsv("meta/lates-lgs-named.txt")
chroms
```

Following https://baylab.github.io/MarineGenomics/week-7-fst-and-outlier-analysis.html#using-fst-to-find-outliers

using outputs/105/renamed-05.vcf.gz, this is the file  of called snps from all our samples in the study

```{sh, eval=FALSE}
bcftools view outputs/105/renamed-05.vcf.gz -S bamlists/38.names | bcftools +fill-tags -- -t MAF,F_MISSING | bcftools view -i 'F_MISSING < 0.05' | bcftools view -i 'INFO/MAF > 0.05' > outputs/115/filt.vcf
bgzip outputs/115/filt.vcf
tabix outputs/115/filt.vcf.gz
bcftools view -Ov -r NC_066833.1,NC_066856.1 outputs/115/filt.vcf.gz > outputs/115/twochroms.vcf
```

956959 snps
```{r, eval=FALSE}
#vcf.path<-"outputs/115/twochroms.vcf"
vcf.path<-"outputs/115/filt.vcf.gz"

data <- read.vcfR(vcf.path)
save(data, file="outputs/115/data.rda")
```

```{r,eval=FALSE}
geno <- extract.gt(data)
save(geno, file="outputs/115/geno.rda")
dim(geno)
head(geno[,1:10])

```

```{r}
load("outputs/115/data.rda")
load("outputs/115/geno.rda")
```


```{r}
G <- geno  #we are doing this because we will be running a lot of different things with G, and if we mess up we want to be able to go back to geno

G[geno %in% c("0/0")] <- 0
G[geno  %in% c("0/1")] <- 1
G[geno %in% c("1/1")] <- 2
G[is.na(G)] <- 9
tG <- t(G)
dim(tG)
subgen<-tG
```

Don't need to subset
```{r}
fst <- MakeDiploidFSTMat(subgen,locusNames=1:ncol(subgen),popNames=m38$NewPhenotype)
```

```{r}
head(fst)
hist(fst$FST,breaks=50)

```
```{r}
summary(fst$FST) #highest FST is higher than the mean (which is a good sign)
```

Looking for low H2, high FST loci
````{r}
plot(fst$He, fst$FST)

```

Once we’ve calculated Fst between the two populations for each SNP individually, we want to determine whether some SNPs are statistical outliers - that is, more differentiated than we would expect. OutFLANK does this by fitting a Chi-Squared distribution to the data and looking to see if the tails of the Chi-Squared distribution have more SNPs than expected:

Suggested values are 0.05 not 0.01 for right and left Trim increasing snps from 28 to 30
Hm is 0.05, moving to 0.10, snps from 28 to 30
Number of samples - mumber of spatial locaitons? setting to two

qthreshold of 0.05 and 0.01 return the same results.

```{r}
OF <- OutFLANK(fst,LeftTrimFraction=0.01,RightTrimFraction=0.01,
         Hmin=0.1,NumberOfSamples=2,qthreshold=0.05)
OutFLANKResultsPlotter(OF,withOutliers=T,
                       NoCorr=T,Hmin=0.1,binwidth=0.005,
                       Zoom=F,RightZoomFraction=0.05,titletext=NULL)
```

```{r}
P1 <- pOutlierFinderChiSqNoCorr(fst,Fstbar=OF$FSTNoCorrbar,
                                dfInferred=OF$dfInferred,qthreshold=0.05,Hmin=0.1)
outliers <- P1$OutlierFlag==TRUE #which of the SNPs are outliers?
table(outliers)

```

887311

```{r}
plot(P1$LocusName,P1$FST,xlab="Position",ylab="FST",col=rgb(0,0,0,alpha=0.1))
points(P1$LocusName[outliers],P1$FST[outliers],col="magenta")
```

Making a plot based on chroms
```{r}
snpmeta<-data@fix %>% as_tibble() %>% mutate(LocusName=1:n())
p2<-P1 %>% left_join(snpmeta) %>% arrange(CHROM)
outs<-p2 %>% filter(OutlierFlag==TRUE)

write_csv(p2, file="outputs/115/outflank.csv")
write_csv(outs, file="outputs/115/outflank-outs.csv")
labels<-p2 %>% group_by(CHROM) %>% summarize(Min=min(LocusName), Max=max(LocusName)) %>% mutate(Mid=(Min+Max)/2) %>% left_join(chroms, by=c("CHROM"="Chromosome")) %>% mutate(Steps=lag(Max))
```

```{R}
nb.cols <- nrow(labels)
mycolors <- colorRampPalette(brewer.pal(8, "Set1"))(nb.cols)

```

```{r}
ggplot(p2 %>% filter(FST >0)) +
  geom_rect(data=labels, aes(xmin=Min, xmax=Max), ymin=0, ymax=0.55, fill=mycolors, alpha=0.25) +
  geom_point(aes(x=LocusName, y=FST, color=CHROM), alpha=0.5) +
  geom_point(data=outs, aes(x=LocusName, y=FST), fill="black") +
  theme_bw() +
  #geom_hline(yintercept = min(outs$FST), alpha=0.5) +
  #geom_vline(xintercept = labels$Steps) +
  theme(panel.grid=element_blank()) +
  scale_x_continuous(breaks=labels$Mid, labels = labels$Chrom) +
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1)) +
  xlab("") +
  ylab("Fst\n") +
  theme(legend.position = "none") +
  scale_color_manual(values=mycolors) +
  theme(axis.title.y = element_text(face="bold", size=14))


ggsave("outputs/115/fst-outlier-plot.jpeg", width=8, height=5)
ggsave("outputs/115/fst-outlier-plot.pdf", width=8, height=5)

```


```{r}
outs %>% left_join(chroms, by=c("CHROM"="Chromosome")) %>% select(CHROM, Chrom, POS, FST) %>% group_by(CHROM, Chrom) %>% summarize(Count=n()) %>% arrange(-Count)
```

```{r}
outs %>% left_join(chroms, by=c("CHROM"="Chromosome")) %>% select(CHROM, Chrom, POS, FST) %>% group_by(CHROM, Chrom) %>% mutate(Count=n()) %>% arrange(-Count, -FST)
```
```{r}
outs %>% arrange(CHROM, POS) %>% select(CHROM, POS, FST, qvalues, INFO)
```

```{r}
outs %>% arrange(CHROM, POS) %>% select(CHROM, POS, FST, qvalues, INFO) %>% filter(CHROM=="NC_066856.1") %>% arrange(-FST) %>% write_csv("outputs/115/top-outs-24.csv")
```
# GWAS

```{sh, eval=FALSE}
srun -t 72:00:00 -p bmh --mem=32G --nodes=1 --ntasks-per-node=1 --cpus-per-task=12 $HOME/angsd/angsd -P 12 \
 -doAsso 1 -yBin $HOME/barra-lh/bamlists/38.pheno -GL 1 -minInd 36 -minMaf 0.10 \
 -minMapQ 20 -minQ 20 \
 -doMajorMinor 1 -doMaf 1 -SNP_pval 1e-6 -rf meta/lates-lgs.txt -out outputs/115/combined-asso-95 \
 -bam $HOME/barra-lh/bamlists/38.bamlist  > outputs/115/combined-asso.out 2> outputs/115/combined-asso.err &
```
