---
title: "116-clustering-redo"
output: html_document
date: "2026-04-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(tidyverse)
library(pcadapt)
library(grid)
library(viridis)
library(ggplotify)
library(pheatmap)
library(ggrepel)
library(ggpubr)
library(grid)
library(raster)
library(ozmaps)
library(ggrepel)
library(ggspatial)
```

We have to cluster some stuff!

```{r}
chroms<-read_tsv("meta/lates-lgs-named.txt")
qld<-read_csv("meta/qld-fish.csv") %>% mutate(NewPhenotype=Phenotype) %>% filter(Group!="Mainstream")
mq<-read_csv("~/github/barra-lh/meta/meta44.csv") %>% mutate(Group="Burdekin") %>% bind_rows(qld) %>% filter(Coverage > 8)
mq$Group<-gsub("Burdekin","Burdekin Region",mq$Group)
mq$Group<-gsub("FitzroyQLD","Fitzroy River",mq$Group)
mq$Group<-factor(mq$Group, levels=c("Daly River","Roper River","Burdekin Region","Fitzroy River"))
mq<-mq %>% mutate(PlotPheno=ifelse(Group == "Burdekin Region",NewPhenotype, ifelse(Group=="Daly River", "Daly River Migratory","Unknown")))
mq$PlotPheno<-gsub("SFF","Burdekin Region Migratory",mq$PlotPheno)
mq$PlotPheno<-gsub("SSS","Burdekin Region Resident",mq$PlotPheno)

```

## Outlier clustering


```{r}
meta<-mq
outliers<-read_csv("outputs/115/outflank-outs.csv") %>% arrange(-FST) #%>% # head(n=5)
 # filter(CHROM=="NC_066856.1") %>% head(n=5) 
  #filter(POS %in% c("3001384","2992178","2992188","2993681","2995063"))

ddf<-read_tsv("outputs/105/snps-wgs-05-glf.geno.gz", col_names = c("FALSE"))  %>% dplyr::select(-X72) %>% mutate(ID=paste0(`FALSE`,"_",X2)) %>%
  relocate(ID) %>% left_join(chroms, by=c(`FALSE` = "Chromosome")) %>% relocate(Chrom) %>% mutate(ID2=paste0(Chrom," ",X2)) 
#%>% filter(`FALSE`=="NC_066856.1") 
ddf[ddf==-1]<-NA
ddf2<-ddf %>% filter(ID %in% outliers$ID) 
m<-ddf2 %>% dplyr::select(-ID, -ID2,-Chrom,-`FALSE`, -X2) %>% as.matrix()
annot<-data.frame(sample=as.character(colnames(m))) %>% column_to_rownames("sample")
```




```{r}
annot$Phenotype<-meta$PlotPheno
annot$`Sampling Location`<-meta$Group
annot_colors=list(Phenotype=c("Burdekin Region Resident"="blue","Burdekin Region Migratory"="red","Daly River Migratory"="deeppink4",Unknown="grey75"),"Sampling Location"=c("Daly River"="pink","Roper River"="purple","Burdekin Region"="darkorange1", "Fitzroy River"="darkgoldenrod1"))

```

 color=colorRampPalette(c("#000004FF","#B63679FF","#FCFDBFFF"))))
 
```{r}
pheatmap(m,annotation_col = annot, cluster_rows = TRUE, annotation_colors = annot_colors, color=colorRampPalette(viridis(3, option="cividis", direction=-1, alpha = 0.75))(3),
                   labels_col = meta$specimen_id, labels_row = ddf2$ID2, fontsize_row = 14, fontsize_col = 6)
grid.text("foo", x=-20, rot=90, gp=gpar(fontsize=25))

```


```{r}
as.ggplot(
pheatmap(m, annotation_col = annot, cluster_rows = TRUE,  annotation_colors = annot_colors, color=colorRampPalette(viridis(3, option="cividis", direction=-1, alpha = 0.75))(3),
                   labels_col = meta$specimen_id, labels_row = ddf2$ID2, fontsize_row = 8, fontsize_col = 6,
         angle_col = 315)) +
         
#  ggtitle("Heatmap of associated barra variants on Lca24\n") +
  labs(caption="\nIndividual") +
  theme(plot.title = element_text(hjust=0.5)) +
  theme(plot.caption = element_text(size=14, hjust=0.4, face="bold")) 

ggsave("outputs/116/top-associated-variants.jpeg", width=14, height=6)
ggsave("outputs/116/top-associated-variants.pdf", width=14, height=6)

```



> mq %>% group_by(NewPhenotype) %>% summarize(Count=n())
# A tibble: 3 × 2
  NewPhenotype Count
  <chr>        <int>
1 SFF             22
2 SSS             22
3 UNK             25

Cluster 1: 21 fish
17 SSS
2 SFF
2 Fitz

Cluster 2:
5  SSS
20 SFF
5 Roper
8 Daly
10 Fitzroy

Clustering accuracy
(17+20)



clustering_method = "ward.D"
```{r}
as.ggplot(
pheatmap(m, annotation_col = annot, cluster_rows = TRUE,  annotation_colors = annot_colors, color=colorRampPalette(viridis(3, option="cividis", direction=-1, alpha = 0.75))(3),
                   labels_col = meta$specimen_id, labels_row = ddf2$ID2, fontsize_row = 8, fontsize_col = 6, clustering_method = "ward.D",
         angle_col = 315)) +
         
#  ggtitle("Heatmap of associated barra variants on Lca24\n") +
  labs(caption="\nIndividual") +
  theme(plot.title = element_text(hjust=0.5)) +
  theme(plot.caption = element_text(size=14, hjust=0.4, face="bold")) 

ggsave("outputs/116/top-associated-variants-ward.jpeg", width=12, height=6)
ggsave("outputs/116/top-associated-variants-ward.pdf", width=12, height=6)

```

# Look at frequency of outliers across pops?

 outputs/105/renamed-05.vcf.gz 

```{r}
mq %>% filter(Group=="Fitzroy River") %>% select(specimen_id) %>% write_tsv("outputs/116/fitz.tsv", col_names = FALSE)
mq %>% filter(Group=="Daly River") %>% select(specimen_id) %>% write_tsv("outputs/116/daly.tsv", col_names = FALSE)
mq %>% filter(Group=="Roper River") %>% select(specimen_id) %>% write_tsv("outputs/116/roper.tsv", col_names = FALSE)
bm %>% filter(PlotPheno=="Burdekin Region Migratory") %>% select(specimen_id) %>% write_tsv("outputs/116/mig.tsv", col_names = FALSE)
bm %>% filter(PlotPheno=="Burdekin Region Resident") %>% select(specimen_id) %>% write_tsv("outputs/116/res.tsv", col_names = FALSE)

bm %>% filter(Group=="Burdekin Region") %>% select(specimen_id) %>% write_tsv("outputs/116/burd.tsv", col_names = FALSE)

read_csv("outputs/115/outflank-outs.csv") %>% select(CHROM, POS) %>% write_tsv("outputs/116/outs.tsv", col_names = FALSE)
```

```{sh, eval=FALSE}
bcftools view outputs/105/renamed-05.vcf.gz -R outputs/116/outs.tsv | bcftools view -S outputs/116/fitz.tsv | bcftools +fill-tags -- -t MAF,F_MISSING | grep -v "#" | cut -f 1-8 > outputs/116/fitz.dat

bcftools view outputs/105/renamed-05.vcf.gz -R outputs/116/outs.tsv | bcftools view -S outputs/116/daly.tsv | bcftools +fill-tags -- -t MAF,F_MISSING | grep -v "#" | cut -f 1-8 > outputs/116/daly.dat

bcftools view outputs/105/renamed-05.vcf.gz -R outputs/116/outs.tsv | bcftools view -S outputs/116/roper.tsv | bcftools +fill-tags -- -t MAF,F_MISSING | grep -v "#" | cut -f 1-8 > outputs/116/roper.dat

bcftools view outputs/105/renamed-05.vcf.gz -R outputs/116/outs.tsv | bcftools view -S outputs/116/mig.tsv | bcftools +fill-tags -- -t MAF,F_MISSING | grep -v "#" | cut -f 1-8 > outputs/116/mig.dat

bcftools view outputs/105/renamed-05.vcf.gz -R outputs/116/outs.tsv | bcftools view -S outputs/116/res.tsv | bcftools +fill-tags -- -t MAF,F_MISSING | grep -v "#" | cut -f 1-8 > outputs/116/res.dat

bcftools view outputs/105/renamed-05.vcf.gz -R outputs/116/outs.tsv | bcftools view -S outputs/116/burd.tsv | bcftools +fill-tags -- -t MAF,F_MISSING | grep -v "#" | cut -f 1-8 > outputs/116/burd.dat
```


```{r}
f<-read_tsv("outputs/116/fitz.dat", col_names = FALSE) %>% separate(X8, into=c("PR","AC","AN","Missing","MAF"), sep=";") %>% mutate(Group="Fitzroy River") %>% mutate(Index=1:n())
f$MAF<-gsub("MAF=","",f$MAF)
f$MAF<-as.numeric(f$MAF)


d<-read_tsv("outputs/116/daly.dat", col_names = FALSE) %>% separate(X8, into=c("PR","AC","AN","Missing","MAF"), sep=";")  %>% mutate(Group="Daly River")  %>% mutate(Index=1:n())
d$MAF<-gsub("MAF=","",d$MAF)
d$MAF<-as.numeric(d$MAF)

r<-read_tsv("outputs/116/roper.dat", col_names = FALSE) %>% separate(X8, into=c("PR","AC","AN","Missing","MAF"), sep=";")  %>% mutate(Group="Roper River")  %>% mutate(Index=1:n())
r$MAF<-gsub("MAF=","",r$MAF)
r$MAF<-as.numeric(r$MAF)

m<-read_tsv("outputs/116/mig.dat", col_names = FALSE) %>% separate(X8, into=c("PR","AC","AN","Missing","MAF"), sep=";")  %>% mutate(Group="Burdekin Region Migratory")  %>% mutate(Index=1:n())
m$MAF<-gsub("MAF=","",m$MAF)
m$MAF<-as.numeric(m$MAF)

res<-read_tsv("outputs/116/res.dat", col_names = FALSE) %>% separate(X8, into=c("PR","AC","AN","Missing","MAF"), sep=";")  %>% mutate(Group="Burdekin Region Resident")  %>% mutate(Index=1:n())
res$MAF<-gsub("MAF=","",res$MAF)
res$MAF<-as.numeric(res$MAF)

b<-read_tsv("outputs/116/burd.dat", col_names = FALSE) %>% separate(X8, into=c("PR","AC","AN","Missing","MAF"), sep=";")  %>% mutate(Group="Burdekin Region")  %>% mutate(Index=1:n())
b$MAF<-gsub("MAF=","",b$MAF)
b$MAF<-as.numeric(b$MAF)
```

Checking for same ref/alts
```{r}
select(f, X3, X4, X5, Index) %>% left_join(b)
```


```{r}
ddf<-select(f,Index, X1, X2, MAF, Group) %>% bind_rows(select(d, Index, X1, X2, MAF, Group)) %>% bind_rows(select(r, Index, X1, X2, MAF, Group))  %>% bind_rows(select(b, Index, X1, X2, MAF, Group)) %>% bind_rows(select(m, Index, X1, X2, MAF, Group))  %>% bind_rows(select(res, Index, X1, X2, MAF, Group))  %>% left_join(chroms, by=c("X1"="Chromosome"))
ddf$Group<-factor(ddf$Group, levels=c("Daly River","Roper River","Burdekin Region","Burdekin Region Resident","Burdekin Region Migratory","Fitzroy River"))

ls<-ddf %>% select(X1, X2, Index) %>% unique() %>% left_join(chroms, by=c("X1"="Chromosome"))
ls
```

```{r}
ggplot(ddf) +
  geom_line(aes(x=Index, y=MAF, group=Group, color=Group)) +
  geom_point(aes(x=Index, y=MAF, color=Group)) +
 # facet_grid(Group~.) +
  theme_bw() +
  theme(panel.grid=element_blank()) +
  scale_color_viridis_d(option="H") +
  #theme(legend.position = "none") +
  scale_x_continuous(name="\nGenomic Location", breaks=ls$Index, labels=paste0(ls$Chrom, " ",ls$X2)) +
  theme(axis.text.x = element_text(angle=45, hjust=1)) +
  ylab("MAF\n")
  

ggsave("outputs/116/MAF-plot.jpeg", width=6, height=10)
ggsave("outputs/116/MAF-plot.pdf", width=6, height=10)

```


Cluster based on allele freqs...

```{r}
d2<-ddf %>% relocate(Chrom) %>% pivot_wider(names_from = Group, values_from = MAF) %>% mutate("Genomic Position"=paste0(Chrom, " ", X2))

write_csv(d2, file="outputs/116/allele-freqs.csv")
mat<-d2 %>%   dplyr::select(-Chrom, -Index,-X1,-X2,-"Genomic Position") %>% as.matrix()
  
annotmat<-data.frame(sample=as.character(colnames(mat))) %>% column_to_rownames("sample")
annotmat$Phenotype<-c("Unknown","Daly River Migratory","Unknown","Unknown","Burdekin Region Migratory","Burdekin Region Resident")

annotmat$`Sampling Location`<-c("Fitzroy River","Daly River","Roper River","Burdekin Region","Burdekin Region","Burdekin Region")
annot_colors=list(Phenotype=c("Burdekin Region Resident"="blue","Burdekin Region Migratory"="red","Daly River Migratory"="deeppink4",Unknown="grey75"),"Sampling Location"=c("Daly River"="pink","Roper River"="purple","Burdekin Region"="darkorange1", "Fitzroy River"="darkgoldenrod1"))

```

```{r}
as.ggplot(pheatmap(mat, cluster_rows = TRUE, color=colorRampPalette(viridis(10, option="mako", direction=1, alpha = 0.75))(10), labels_row = d2$`Genomic Position`, annotation_col = annotmat,  annotation_colors = annot_colors)
)

ggsave("outputs/116/allele-freq-clust.pdf", width=10, height=6)

```

Clustering accuracy?
44 fish

```{r}
(17+20)/44
(2+5)/44
```

```