---
title: "117-pike-dapc"
output: html_document
date: "2026-04-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```


```{r}
library(adegenet)
library(dartR)
library(vcfR)
library(tidyverse)
```

Separating fish with DAPC (using pruned loci at first, then we can use unpruned?)
```{sh, eval=FALSE}

grep CR vcf-meta.tsv | cut -f 1 > chat.tsv
bcftools view -S chat.tsv Eluc.variants.GATK.iteration.2.b.Full.SNP.GATK_HF_removed.minQ20.mmdp10.mxmdp60.mmc10.mac1.homsumfilt.new.samp.names.recode.vcf.gz | bcftools +fill-tags -- -t MAF,F_MISSING | bcftools view -i 'F_MISSING < 0.05' | bcftools view -i 'INFO/MAF > 0.10' > chat.vcf

grep "S" vcf-meta.tsv  > s-meta.tsv
grep "S" vcf-meta.tsv | grep -v Sample | cut -f 1 > s-samples.txt
bcftools view -S s-samples.txt Eluc.variants.GATK.iteration.2.b.Full.SNP.GATK_HF_removed.minQ20.mmdp10.mxmdp60.mmc10.mac1.homsumfilt.new.samp.names.recode.vcf.gz | bcftools +fill-tags -- -t MAF,F_MISSING | bcftools view -i 'F_MISSING < 0.05' | bcftools view -i 'INFO/MAF > 0.10' > s.vcf

bcftools view -S east.tsv Eluc.variants.GATK.iteration.2.b.Full.SNP.GATK_HF_removed.minQ20.mmdp10.mxmdp60.mmc10.mac1.homsumfilt.new.samp.names.recode.vcf.gz | bcftools +fill-tags -- -t MAF,F_MISSING | bcftools view -i 'F_MISSING < 0.05' | bcftools view -i 'INFO/MAF > 0.10' > east.vcf

#thinned version in outputs/117

bgzip filtered.vcf
tabix filtered.vcf.gz.

gunzip -c filtered.vcf.gz | grep "#" > header.txt 
gunzip -c filtered.vcf.gz | grep -v "#" | awk 'BEGIN{i=0}{i++;if (i%100==0) print}' > t100.vcf
gunzip -c filtered.vcf.gz | grep -v "#" | awk 'BEGIN{i=0}{i++;if (i%50==0) print}' > t50.vcf
gunzip -c filtered.vcf.gz | grep -v "#" | awk 'BEGIN{i=0}{i++;if (i%20==0) print}' > t20.vcf
gunzip -c filtered.vcf.gz | grep -v "#" | awk 'BEGIN{i=0}{i++;if (i%10==0) print}' > t10.vcf

cat header.txt t20.vcf > thinned20.vcf
cat header.txt t10.vcf > thinned10.vcf

bcftools +fill-tags thinned10.vcf.gz -- -t MAF,F_MISSING  | bcftools view -i 'INFO/MAF > 0.10' > filtered.vcf
bgzip filtered.vcf
tabix filtered.vcf.gz
bgzip thinned20.vcf
bgzip thinned10.vcf

tabix thinned20.vcf.gz
tabix thinned10.vcf.gz
```

112794 variants in t10.vcf.gz

```{r, eval=FALSE}
d<-read.vcfR("/Users/mac/Dropbox/ak-pike-genomics/christensen-et-al/Eluc.variants.GATK.iteration.2.b.Full.SNP.GATK_HF_removed.minQ20.mmdp10.mxmdp60.mmc10.mac1.homsumfilt.new.samp.names.recode.vcf.gz")

obj<-vcfR2genlight(d)
save(obj, file="/Users/mac/Dropbox/ak-pike-genomics/christensen-et-al/obj.rda")
```

```{r, eval=TRUE}
d<-read.vcfR("/Users/mac/Dropbox/ak-pike-genomics/christensen-et-al/chat.vcf")

obj<-vcfR2genlight(d)
```

```{r}
meta<-read_tsv("/Users/mac/Dropbox/ak-pike-genomics/christensen-et-al/vcf-meta.tsv")
meta
```

```{r}
obj@ind.names
```


# DAPC with sex, Chat

```{r}
chat<-meta %>% filter(Pop=="CR")
obj2<-obj[obj@ind.names %in% chat$Sample,]
```

```{r}
dapc<-dapc(obj2, as.factor(chat$Sex), n.pca = 40, n.da = 1)
```

```{r}
scatter(dapc)
```

Let's make a plot of assignments


```{r}
chat$Assign<-dapc$assign
male<-dapc$posterior[,1] %>% as_tibble() %>% rename(F=value)
female<-dapc$posterior[,2] %>% as_tibble() %>% rename(M=value)

bm<-bind_cols(chat,male,female)
``` 

```{r}
b2<-bm %>% dplyr::select(Sample, Pop, Sex, Assign, 'F', M ) %>% arrange(Sex,-F) %>% mutate(Index=1:n()) %>% relocate(Index) %>% pivot_longer(6:7, names_to = "AssignedSex", values_to = "Posterior")

labels<-b2 %>% group_by(Sex) %>% summarize(Start=min(Index), Stop=max(Index), Position=mean(Index))

ticks<-b2 %>% select(Index, Sample,Sex) %>% unique()
```


```{r}
ggplot(b2) +
  geom_col(aes(x=Index, y=Posterior, fill=AssignedSex), width=1, alpha=0.75) +
  geom_segment(data=labels, x = labels$Start - 0.5, y=0, xend = labels$Start-0.5, yend=1, alpha=0.9,  size=1) +
  geom_segment(data=labels, x = labels$Stop[length(labels$Stop)]  + 0.5, y=0, 
               xend= labels$Stop[length(labels$Stop)] + 0.5, yend=1,  alpha=0.9,
               size=1) +
  geom_segment(x=0+0.4, xend= labels$Stop[length(labels$Stop)]+0.6, y=1, yend=1, alpha=0.9, size=1) +
  geom_segment(x=0+0.4, xend= labels$Stop[length(labels$Stop)]+0.6, y=0, yend=0, alpha=0.9, size=1) +
  scale_fill_manual(values=c("red","blue"), name="Genetic Assignment", labels=c("Female","Male")) +
  ylab("Posterior Probability of Assignment\n") +
  xlab("\nIndividual") +
  scale_x_continuous(breaks=ticks$Index, labels=ticks$Sample) +
  theme(axis.text.x = element_text(angle=45, hjust=1)) +
  theme(panel.background = element_blank()) +
  theme(axis.text.x = element_text(size=6)) +
  theme(axis.text.y = element_text(size=8)) +
  theme(axis.title.y=element_text(size=12, face="bold")) +
  theme(axis.title.x=element_text(size=12, face="bold")) +
  geom_text(data=labels, aes(x=Position), label=c("Phenotypic Females","Phenotypic Males"), y=1.04, size=3) +
  ylim(0,1.03)
  
ggsave("outputs/117/posterior-assignment-dapc.jpeg", width=9, height=4)
ggsave("outputs/117/posterior-assignment-dapc.pdf", width=9, height=4)

```
Get sites with the most loadings:
```{r}
contrib <- loadingplot(dapc$var.contr, axis=1,
thres=.002, lab.jitter=0)

contribs<-as.tibble(dapc$var.contr)
contribs$Chromosome<-obj2@chromosome
contribs$Position<-obj2@position
contribs<-mutate(contribs, number=1:n())

```
```{r}
topfrac<-top_frac(contribs, n=.0001, LD1)
topfrac %>% arrange(-LD1)

write_csv(topfrac, "outputs/117/loadings.csv")
```

# DAPC with sex, S

```{r, eval=TRUE}
d<-read.vcfR("/Users/mac/Dropbox/ak-pike-genomics/christensen-et-al/s.vcf")

obj<-vcfR2genlight(d)
```

```{r}
s<-meta %>% filter(Pop=="S")
obj2<-obj[obj@ind.names %in% s$Sample,]
```

```{r}
dapc<-dapc(obj2, as.factor(s$Sex), n.pca = 40, n.da = 1)
```

```{r}
scatter(dapc)
```

Let's make a plot of assignments


```{r}
s$Assign<-dapc$assign
male<-dapc$posterior[,1] %>% as_tibble() %>% rename(F=value)
female<-dapc$posterior[,2] %>% as_tibble() %>% rename(M=value)

bm<-bind_cols(s,male,female)
``` 

```{r}
b2<-bm %>% dplyr::select(Sample, Pop, Sex, Assign, 'F', M ) %>% arrange(Sex,-F) %>% mutate(Index=1:n()) %>% relocate(Index) %>% pivot_longer(6:7, names_to = "AssignedSex", values_to = "Posterior")

labels<-b2 %>% group_by(Sex) %>% summarize(Start=min(Index), Stop=max(Index), Position=mean(Index))

ticks<-b2 %>% select(Index, Sample,Sex) %>% unique()
```


```{r}
ggplot(b2) +
  geom_col(aes(x=Index, y=Posterior, fill=AssignedSex), width=1, alpha=0.75) +
  geom_segment(data=labels, x = labels$Start - 0.5, y=0, xend = labels$Start-0.5, yend=1, alpha=0.9,  size=1) +
  geom_segment(data=labels, x = labels$Stop[length(labels$Stop)]  + 0.5, y=0, 
               xend= labels$Stop[length(labels$Stop)] + 0.5, yend=1,  alpha=0.9,
               size=1) +
  geom_segment(x=0+0.4, xend= labels$Stop[length(labels$Stop)]+0.6, y=1, yend=1, alpha=0.9, size=1) +
  geom_segment(x=0+0.4, xend= labels$Stop[length(labels$Stop)]+0.6, y=0, yend=0, alpha=0.9, size=1) +
  scale_fill_manual(values=c("red","blue"), name="Genetic Assignment", labels=c("Female","Male")) +
  ylab("Posterior Probability of Assignment\n") +
  xlab("\nIndividual") +
  scale_x_continuous(breaks=ticks$Index, labels=ticks$Sample) +
  theme(axis.text.x = element_text(angle=45, hjust=1)) +
  theme(panel.background = element_blank()) +
  theme(axis.text.x = element_text(size=6)) +
  theme(axis.text.y = element_text(size=8)) +
  theme(axis.title.y=element_text(size=12, face="bold")) +
  theme(axis.title.x=element_text(size=12, face="bold")) +
  geom_text(data=labels, aes(x=Position), label=c("Phenotypic Females","Phenotypic Males"), y=1.04, size=3) +
  ylim(0,1.03)
  
ggsave("outputs/117/posterior-assignment-dapc-s.jpeg", width=9, height=4)
ggsave("outputs/117/posterior-assignment-dap-s.pdf", width=9, height=4)

```
Get sites with the most loadings:
```{r}
contrib <- loadingplot(dapc$var.contr, axis=1,
thres=.002, lab.jitter=0)

contribs<-as.tibble(dapc$var.contr)
contribs$Chromosome<-obj2@chromosome
contribs$Position<-obj2@position
contribs<-mutate(contribs, number=1:n())

```
```{r}
topfrac<-top_frac(contribs, n=.001, LD1)
topfrac %>% arrange(-LD1)

write_csv(topfrac, "outputs/117/loadings-s.csv")
```

# DAPC with sex, SLA, WHI HCK

```{r, eval=TRUE}
d<-read.vcfR("/Users/mac/Dropbox/ak-pike-genomics/christensen-et-al/east.vcf")

obj<-vcfR2genlight(d)
```

```{r}
s<-meta %>% filter(Pop %in% c("S","NJ","Mb"))
obj2<-obj[obj@ind.names %in% s$Sample,]
```

```{r}
dapc<-dapc(obj2, as.factor(s$Sex), n.pca = 40, n.da = 1)
```

```{r}
scatter(dapc)
```

Let's make a plot of assignments


```{r}
s$Assign<-dapc$assign
male<-dapc$posterior[,1] %>% as_tibble() %>% rename(F=value)
female<-dapc$posterior[,2] %>% as_tibble() %>% rename(M=value)

bm<-bind_cols(s,male,female)
``` 

```{r}
b2<-bm %>% dplyr::select(Sample, Pop, Sex, Assign, 'F', M ) %>% arrange(Sex,-F) %>% mutate(Index=1:n()) %>% relocate(Index) %>% pivot_longer(6:7, names_to = "AssignedSex", values_to = "Posterior")

labels<-b2 %>% group_by(Sex) %>% summarize(Start=min(Index), Stop=max(Index), Position=mean(Index))

ticks<-b2 %>% select(Index, Sample,Sex) %>% unique()
```


```{r}
ggplot(b2) +
  geom_col(aes(x=Index, y=Posterior, fill=AssignedSex), width=1, alpha=0.75) +
  geom_segment(data=labels, x = labels$Start - 0.5, y=0, xend = labels$Start-0.5, yend=1, alpha=0.9,  size=1) +
  geom_segment(data=labels, x = labels$Stop[length(labels$Stop)]  + 0.5, y=0, 
               xend= labels$Stop[length(labels$Stop)] + 0.5, yend=1,  alpha=0.9,
               size=1) +
  geom_segment(x=0+0.4, xend= labels$Stop[length(labels$Stop)]+0.6, y=1, yend=1, alpha=0.9, size=1) +
  geom_segment(x=0+0.4, xend= labels$Stop[length(labels$Stop)]+0.6, y=0, yend=0, alpha=0.9, size=1) +
  scale_fill_manual(values=c("red","blue"), name="Genetic Assignment", labels=c("Female","Male")) +
  ylab("Posterior Probability of Assignment\n") +
  xlab("\nIndividual") +
  scale_x_continuous(breaks=ticks$Index, labels=ticks$Sample) +
  theme(axis.text.x = element_text(angle=45, hjust=1)) +
  theme(panel.background = element_blank()) +
  theme(axis.text.x = element_text(size=6)) +
  theme(axis.text.y = element_text(size=8)) +
  theme(axis.title.y=element_text(size=12, face="bold")) +
  theme(axis.title.x=element_text(size=12, face="bold")) +
  geom_text(data=labels, aes(x=Position), label=c("Phenotypic Females","Phenotypic Males"), y=1.04, size=3) +
  ylim(0,1.03)
  
ggsave("outputs/117/posterior-assignment-dapc-east.jpeg", width=9, height=4)
ggsave("outputs/117/posterior-assignment-dap-east.pdf", width=9, height=4)

```
Get sites with the most loadings:
```{r}
contrib <- loadingplot(dapc$var.contr, axis=1,
thres=.002, lab.jitter=0)

contribs<-as.tibble(dapc$var.contr)
contribs$Chromosome<-obj2@chromosome
contribs$Position<-obj2@position
contribs<-mutate(contribs, number=1:n())

```
```{r}
topfrac<-top_frac(contribs, n=.0001, LD1)
topfrac %>% arrange(-LD1)

write_csv(topfrac, "outputs/117/loadings-east.csv")
```


Train on half and then assign.

```{r}
east<-meta %>% filter(Pop %in% c("S","NJ","Mb")) 
train<-east %>% group_by(Sex) %>% slice_sample(n=5) %>% ungroup()
test<-east %>% filter(!Sample %in% sub$Sample)
```

```{r}
ot<-obj[obj@ind.names %in% train$Sample,]
otest<-obj[obj@ind.names %in% test$Sample,]
```

```{r}
dapc<-dapc(ot, as.factor(train$Sex), n.pca = 9, n.da = 1)
```

```{r}
scatter(dapc)
```

Now to assign

# predict.dapc
```{r}
assign<-predict.dapc(object = dapc, obj, dapc$prior)
```

```{r}
post<-assign$posterior %>% as_tibble()
post<-bind_cols(east,post) %>% arrange(Pop, Sex) %>% mutate(Index=1:n())
post<-post %>% select(Index, Pop, Sample, Sex, "F",M) %>% pivot_longer(5:6, names_to = "Predicted", values_to = "Posterior")


labels<-post %>% group_by(Pop, Sex) %>% summarize(Start=min(Index), Stop=max(Index), Position=mean(Index))
ticks<-post %>% select(Index, Sample,Sex) %>% unique()

```

```{r}
ggplot(post) +
  geom_col(aes(x=Index, y=Posterior, fill=Predicted), width = 1, alpha=0.75) +
  geom_segment(data=labels, x = labels$Start - 0.5, y=0, xend = labels$Start-0.5, yend=1, alpha=0.9,  size=2) +
  geom_segment(data=labels, x = labels$Stop[length(labels$Stop)]  + 0.5, y=0, 
               xend= labels$Stop[length(labels$Stop)] + 0.5, yend=1,  alpha=0.9,
               size=2) +
  geom_segment(x=0+0.4, xend= labels$Stop[length(labels$Stop)]+0.6, y=1, yend=1, alpha=0.9, size=2) +
  geom_segment(x=0+0.4, xend= labels$Stop[length(labels$Stop)]+0.6, y=0, yend=0, alpha=0.9, size=2) +
   theme_bw() +
  theme(panel.background = element_blank()) +
  theme(panel.grid = element_blank()) +
  scale_fill_manual(values=c("red","blue"), name="Genetic Assignment", labels=c("Female","Male")) +
  ylab("Posterior Probability of Assignment\n") +
  xlab("\nIndividual") +
  scale_x_continuous(breaks=ticks$Index, labels=ticks$Sample) +
  theme(axis.text.x = element_text(angle=45, hjust=1)) +
  theme(axis.text.x = element_text(size=6)) +
  theme(axis.text.y = element_text(size=8)) +
  theme(axis.title.y=element_text(size=12, face="bold")) +
  theme(axis.title.x=element_text(size=12, face="bold"))  +
  geom_text(data=labels, aes(x=Position), label=labels$Sex, y=1.03, size=3, angle=0) +
  ylim(0,1.03)

ggsave("outputs/117/dapc-assignments-east.pdf", width=12, height = 6)
```


## LD

Have thinned snps, need to separate

```{sh, eval=FALSE}

cat outputs/117/pike-lgs.txt  | while read line; do bcftools view -Ov -r $line outputs/117/thinned10.vcf.gz  > outputs/117/vcfs/$line.vcf; done
cat outputs/117/pike-lgs.txt  | while read line; do bcftools view -Ov -r $line outputs/117/filtered.vcf.gz  > outputs/117/vcfs/$line.vcf; done


for f in *.vcf; do plink --vcf $f --r2 inter-chr --ld-window-r2 0.3 --out `basename $f vcf`ldf --allow-extra-chr --double-id; done;
```

```{r}

files<-list.files("outputs/117/vcfs",pattern = "*.ldf.ld", full.names = TRUE)

plotLd<-function(file) {
  chrom<-gsub("outputs/117/vcfs/","",file)
  chrom<-gsub(".ldf.ld","", chrom)
  lc<-read.delim(file,sep="",stringsAsFactors=FALSE) %>% as_tibble() %>% dplyr::arrange(R2) %>%  filter(R2 >0.8)

  ggplot(lc) +
  geom_point(aes(x=BP_A, y=BP_B, color=R2), alpha=0.5) +
  scale_color_gradient(low="khaki1", high="red") + 
  ggtitle(paste0(chrom))+
  theme_bw() +
  theme(panel.grid = element_blank())
ggsave(paste0("outputs/117/chrom-ld/",chrom,"-ld.pdf"))
}

lapply(files, plotLd)
```

```{r}

files<-list.files("outputs/117/vcfs",pattern = "*.vcf", full.names = TRUE)

plotPCA<-function(file) {
  meta<-meta
  chrom<-gsub("outputs/117/vcfs/","",file)
  chrom<-gsub(".vcf","", chrom)
  
  vcf<-read.vcfR(file=file)
  genind<-vcfR2genind(vcf)
  genind@pop<-as.factor(meta$Pop)

gen<-genind
X <- tab(gen, NA.method="mean")

pca1 <- dudi.pca(X,scannf=FALSE,scale=FALSE,nf=3)

df<-pca1$li
df<-df %>% bind_cols(meta)
eig<-pca1$eig/sum(pca1$eig)*100


pc12<-ggplot(df) + 
  geom_point(aes(x=Axis1, y=Axis2, fill=Pop),pch=21, alpha=0.75, cex=2) +
  theme_bw() +
  xlab(paste0("PC1 ",round(eig[1],2), "%")) +
  ylab(paste0("PC2 ",round(eig[2],2), "%")) +
  theme(panel.grid = element_blank()) +
  scale_fill_viridis_d(option="H") +
  guides(fill = guide_legend(override.aes = list(pch = 22))) +
  theme(legend.position = "top") 
 # geom_text_repel(aes(x=Axis1, y=Axis2,label=id))
pc12

ggsave(paste0("outputs/117/chrom-pcs/",chrom,"-pcs.pdf"))
}

lapply(files, plotPCA)
```


## Local PCA

```{sh, eval=FALSE}
cat outputs/117/pike-lgs.txt  | while read line; do bcftools view -Ob -r $line outputs/117/thinned10.vcf.gz > outputs/117/bcf/$line.bcf; done;
for f in outputs/117/bcf/*.bcf; do bcftools index $f; done;
```

```{r}
samples<-meta %>% select(Sample) %>% rename(ID=Sample)

population<-meta %>% select(Pop) %>% rename(population=Pop)

table<-cbind(samples, population)
write.table(table, "outputs/117/bcf/sample_info.tsv", quote = TRUE, row.names = FALSE, sep="\t")

```

./run_lostruct.R -i /Users/mac/github/barra-lh/outputs/117/bcf -t snp -s 20 -m 4 -I /Users/mac/github/barra-lh/outputs/117/bcf/sample_info.tsv -j 117
cp lostruct_results/type_snp_size_20_weights_none_jobid_117/mds_coords.csv ~/github/barra-lh/outputs/117/mds_coords-20.csv

```{r}
mds<-read_csv("outputs/117/mds_coords-20.csv") # 20 snp windows

#make tidy
tidymds<-mds %>% gather(MDS, Value, 3:6)
MDS1<-filter(tidymds, MDS=="MDS1") %>% rename(MDS1=MDS) %>% rename(Value1=Value)
MDS2<-filter(tidymds, MDS=="MDS2") %>% rename(MDS2=MDS) %>% rename(Value2=Value)
MDS3<-filter(tidymds, MDS=="MDS3") %>% rename(MDS3=MDS) %>% rename(Value3=Value)
MDS4<-filter(tidymds, MDS=="MDS4") %>% rename(MDS4=MDS) %>% rename(Value4=Value)
```
```{r}
ggplot(mds)+
  geom_point(aes(x=MDS1, y=MDS2, fill=chrom), pch=21, alpha=0.75) +
  scale_fill_viridis_d(option="magma") +
  theme_bw() +
  theme(legend.position = 'none')


ggplot(mds)+
  geom_point(aes(x=MDS1, y=MDS3, fill=chrom), pch=21, alpha=0.75) +
  scale_fill_viridis_d(option="magma") +
  theme_bw() +
  theme(legend.position = 'none')


ggplot(mds)+
  geom_point(aes(x=MDS1, y=MDS4, fill=chrom), pch=21, alpha=0.75) +
  scale_fill_viridis_d(option="magma") +
  theme_bw() +
  theme(legend.position = 'none')
```

```{r}
p1<-MDS1 %>% mutate(Index=1:n())

out <- boxplot.stats(p1$Value1)$out
out_ind <- which(p1$Value1 %in% c(out))
length(out_ind)
outliers<-p1[out_ind,]
outliers %>% group_by(chrom) %>% summarize(Count=n()) %>% arrange(-Count)
```

```{r}
chroms<-p1 %>% group_by(chrom) %>% mutate(Start=min(Index), Stop=max(Index)) %>% select(chrom,Start,Stop) %>% unique() %>% 
  mutate(Position=round((Start+Stop)/2))

#Make enough colors
nb.cols <- nrow(chroms)
mycolors <- colorRampPalette(brewer.pal(8, "Set1"))(nb.cols)

ggplot(p1) +
  geom_rect(data=chroms, aes(xmin=Start, xmax=Stop, ymin=min(p1$Value1), ymax=max(p1$Value1)), fill=mycolors, alpha=0.25) +
  geom_point(data=p1, aes(x=Index, y=Value1, color=chrom), alpha=0.75, cex=0.5) +
  geom_point(data=outliers, aes(x=Index, y=Value1), color="black", cex=0.5) +
  theme_bw() +
  scale_x_continuous(breaks=chroms$Position, labels=chroms$chrom) +
  theme(axis.text.x = element_text(angle=45, hjust = 1, face="bold", size=6)) +
  theme(legend.title = element_blank()) +
  theme(panel.grid = element_blank()) +
  theme(legend.position = "none") +
  scale_color_manual(values=mycolors) +
  ylab("MDS1") +
  xlab("Chromosome")
```