---
title: "120-sliding-fst-R2"
output: html_document
date: "2026-07-15"
---

````{r}
library(tidyverse)
library(snpR)
library(ggpubr)
```
We can also look at Fst windows


```{r}
m40<-read_csv("meta/m40.csv")

```


```{sh}
bcftools view outputs/104/renamed-05.vcf.gz -S bamlists/40.names | bcftools +fill-tags -- -t MAF,F_MISSING | bcftools view -i 'F_MISSING < 0.05' | bcftools view -i 'INFO/MAF > 0.05' > outputs/120/filt.vcf
bgzip outputs/120/filt.vcf
tabix outputs/120/filt.vcf.gz
bcftools view -Ov -r NC_066833.1,NC_066856.1 outputs/120/filt.vcf.gz > outputs/120/twochroms.vcf
bcftools view -Ov -r NC_066833.1,NC_066847.1,NC_066856.1 outputs/120/filt.vcf.gz > outputs/120/threechroms.vcf

```

Error: C stack usage 21536065 is too close to the limit
```{r, eval=FALSE}

#snps<-import.snpR.data("outputs/120/threechroms.vcf", sample.meta = dplyr::select(m40,NewPhenotype,NewPheno,specimen_id) %>% rename(FishID=specimen_id))

snps<-import.snpR.data("outputs/120/threechroms.vcf", sample.meta = dplyr::select(m40,NewPhenotype,NewPheno,specimen_id) %>% rename(FishID=specimen_id))

snps@snp.meta$CHROM<-gsub("\\.1","",snps@snp.meta$CHROM)
p<-plot_clusters(snps, "NewPhenotype")
p$plots$pca

x<-calc_association(snps, response = "NewPheno", method = "chisq")
y<-get.snpR.stats(x, stats="association")


plot_manhattan(x, "chi_p_NewPheno", chr = "CHROM",
               log.p = TRUE)$plot

ggsave("outputs/120/snp-r-manhattan.jpeg")
```

```{r, eval=FALSE}
y$single %>% mutate(Position=1:n()) %>% group_by(CHROM) %>%  mutate(BHfdr = p.adjust(chi_p_NewPheno, method = "BH")) %>% arrange(BHfdr) %>% write_csv("outputs/120/snpr-gwas-results.csv")
```


sliding window fst

```{r}
fst<-calc_pairwise_fst(snps, facets="CHROM.NewPheno", method="wc")
fsts<-get.snpR.stats(fst, facets = "CHROM.NewPheno", stats = "fst")
```

```{r}
outliers<-fsts$pairwise %>% as_tibble() %>% mutate(p99 = quantile(fst, probs = 0.999, na.rm = TRUE))  %>% filter(fst > p99)

```

```{r}
ggplot(fsts$pairwise) +
  geom_point(aes(x=position, y=fst), alpha=0.5)  +
  geom_point(data=outliers, aes(x=position, y=fst), pch=21, fill="skyblue") +
  facet_wrap(.~CHROM)
```



```{r}
smooth<-calc_smoothed_averages(fst, facets = "CHROM.NewPheno", stats.type="pairwise", sigma=100, step=50, triple_sigma = FALSE) #sigma=50, step=20
smoothf<-get.snpR.stats(smooth, "CHROM.NewPheno", "pairwise.window") 
```
sigma	
numeric. Designates the width of windows in kilobases. Full window size is 6*sigma or sigma kilobases depending on the triple_sigma argument.

step	
numeric or NULL, default default 2*sigma (non-overlapping windows). Designates the number of kilobases between each window centroid. If NULL, windows are centered on each SNP.



```{r}
outs<-smoothf %>% mutate(p99 = quantile(fst, probs = 0.99, na.rm = TRUE)) %>% filter(fst > p99)

ggplot(smoothf) +
  geom_point(aes(x=position, y=fst), alpha=0.5) +
  geom_point(data=outs, aes(x=position, y=fst), pch=21, fill="skyblue") +
  facet_wrap(.~snp.subfacet, scales="free_x") +
  theme_bw() +
  ylab("Fst") +
  xlab("Position")
```


These files are too big for me to run all together.

Split chroms.

```{sh, eval=FALSE}
cat meta/lates-lgs.txt  | while read line; do bcftools view -Ov -r $line outputs/120/filt.vcf.gz  > outputs/119/vcfs/$line.vcf; done

```

```{r, eval=FALSE}
files<-list.files("outputs/119/vcfs",pattern = "*vcf", full.names = TRUE)

getWindows<-function(file) {
  chrom<-gsub("outputs/119/vcfs/","",file)
  chrom<-gsub(".1.vcf","", chrom)
  snps<-import.snpR.data(file, sample.meta = dplyr::select(m40,NewPhenotype,NewPheno,specimen_id) %>% rename(FishID=specimen_id))
  snps@snp.meta$CHROM<-gsub("\\.1","",snps@snp.meta$CHROM)

  fst<-calc_pairwise_fst(snps, facets="CHROM.NewPheno", method="wc")
#  fsts<-get.snpR.stats(fst, facets = "CHROM.NewPheno", stats = "fst")

  smooth<-calc_smoothed_averages(fst, facets = "CHROM.NewPheno", stats.type="pairwise", sigma=100, step=50, triple_sigma = FALSE) #sigma=50, step=20
  smoothf<-get.snpR.stats(smooth, "CHROM.NewPheno", "pairwise.window") 

  write_csv(smoothf, paste0("outputs/119/",chrom,"-windows.csv"))
}

lapply(files, getWindows)

```

```{r}
files<-list.files("outputs/119",pattern = "*windows.csv", full.names = TRUE)

table<- map_df(files, read_csv, .id = "source_file")

chroms<-read_tsv("meta/lates-lgs-named.txt")
chroms$snp.subfacet<-gsub("\\.1$","",chroms$Chromosome)

table<-left_join(table, chroms, by=c("snp.subfacet"="snp.subfacet"))
table
```


```{r}
outs<-table %>% mutate(p99 = quantile(fst, probs = 0.99, na.rm = TRUE)) %>% filter(fst > p99)
outs
```

```{r}
ggplot(table) +
  geom_point(aes(x=position, y=fst), alpha=0.5, size=1)  +
  geom_point(data=outs, aes(x=position, y=fst), pch=21, fill="skyblue") +
  facet_wrap(.~Chrom, ncol=4) +
  ylab("Fst\n") +
  xlab("\nPosition") +
  theme_bw() +
  theme(axis.title = element_text(size=14, face="bold"))

ggsave("outputs/120/sliding-windows.pdf", width=11, height=8)
ggsave("outputs/120/sliding-windows.jpeg", width=11, height=8)

```

Make like GWAS style

```{r}
nb.cols <- nrow(chroms)
mycolors <- colorRampPalette(brewer.pal(8, "Set1"))(nb.cols)
```


```{r}
t2<-table %>% mutate(Index=1:n())
outs2<-t2 %>% mutate(p99 = quantile(fst, probs = 0.999, na.rm = TRUE)) %>% filter(fst > p99)
outs3<-t2 %>% mutate(p99 = quantile(fst, probs = 0.99, na.rm = TRUE)) %>% filter(fst > p99)

labels<-t2 %>% group_by(Chrom) %>% summarize(Min=min(Index), Max=max(Index)) %>% mutate(Mid=(Min+Max)/2) %>% mutate(Steps=lag(Max))

outs3 %>% group_by(Chrom) %>% summarise(Count=n())
```

```{r}
fstb<-ggplot(t2 %>% filter(fst>=0)) +
  geom_rect(data=labels, aes(xmin=Min, xmax=Max), ymin=0, ymax=0.25, fill=mycolors, alpha=0.25) +
  geom_point(aes(x=Index, y=fst, color=Chrom), alpha=0.5) +
  geom_point(data=outs3, aes(x=Index, y=fst), fill="black") +
  theme_bw() +
 # geom_hline(yintercept = min(outs2$fst), alpha=0.5, lty=2) +
  geom_hline(yintercept = min(outs3$fst), alpha=0.5, lty=2) +

  theme(panel.grid=element_blank()) +
  scale_x_continuous(breaks=labels$Mid, labels = labels$Chrom) +
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1)) +
  xlab("\nChromosome") +
  ylab("Fst\n") +
  theme(legend.position = "none") +
  scale_color_manual(values=mycolors) +
  theme(axis.title = element_text(face="bold", size=14))

fstb


```


```{r}
ggarrange(fsta + ggtitle("A") + theme(plot.title = element_text(size=14, face="bold")), 
          fstb + ggtitle("B") + theme(plot.title = element_text(size=14, face="bold")),
          ncol=1)

ggsave("outputs/120/combined-plot.jpeg", width=8, height=8)
ggsave("outputs/120/combined-plot.pdf", width=8, height=8)

```