---
title: "123-lca15"
output: html_document
date: "2026-07-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message=FALSE)
```

```{r}
library(tidyverse)
library(vcfR)
library(adegenet)
library(ggpubr)
```

```{r}
mq<-read_csv("meta/mq.csv")
mq
```

There's two small elevated windows of LD on lca15

```{r}
file<-"outputs/107/vcfs/NC_066847.1.ldf.ld"
  chrom<-gsub("outputs/107/vcfs/","",file)
  chrom<-gsub(".ldf.ld","", chrom)
  lc<-read.delim(file,sep="",stringsAsFactors=FALSE) %>% as_tibble() %>% dplyr::arrange(R2) %>%  filter(R2 >0.2)

chromld<- ggplot(lc) +
  geom_point(aes(x=BP_A, y=BP_B, color=R2), alpha=0.5) +
  scale_color_gradient(low="khaki1", high="red") + 
  ggtitle(paste0(chrom))+
  theme_bw() +
  theme(panel.grid = element_blank()) +
  xlab("SNP Position A") +
  ylab("SNP Position B")

chromld
```
Let's look at these in more detail.

outputs/105/renamed-05.vcf.gz are our unthinned variants. So, let's zoom in on 'regions'

outlier windows are

33630	NC_066847.1	Lca15	13300000	0.12709919	13333630	0.3347389	6.016872e-03	PR	AC=32	
16370	NC_066847.1	Lca15	13350000	0.15687052	13333630	0.3347389	6.016872e-03	PR	AC=32	
13426	NC_066847.1	Lca15	13350000	0.15687052	13363426	0.3131144	9.498987e-03	PR	AC=34	
36574	NC_066847.1	Lca15	13400000	0.13383349	13363426	0.3131144	9.498987e-03	PR	AC=34	
26269	NC_066847.1	Lca15	13350000	0.15687052	13376269	0.3131144	9.498987e-03	PR	AC=34	
23731	NC_066847.1	Lca15	13400000	0.13383349	13376269	0.3131144	9.498987e-03	PR	AC=34	
37316	NC_066847.1	Lca15	17550000	0.05055258	17587316	0.3239579	7.530711e-03	PR	AC=24	
12684	NC_066847.1	Lca15	17600000	0.11329964	17587316	0.3239579	7.530711e-03	PR	AC=24	
17454	NC_066847.1	Lca15	27800000	0.03285632	27782546	0.3415812	6.016872e-03	PR	AC=13	
17410	NC_066847.1	Lca15	27800000	0.03285632	27782590	0.3353902	6.016872e-03	PR	AC=13	
17389	NC_066847.1	Lca15	27800000	0.03285632	27782611	0.3353902	6.016872e-03	PR	AC=13	
16956	NC_066847.1	Lca15	27800000	0.03285632	27783044	0.3353902	6.016872e-03	PR	AC=13	

13250000 - 13450000 (~2e5 bp)
17500000 - 17650000


subreiong three
```{sh, eval=FALSE}
bcftools view -r NC_066847.1:12000000-19000000 outputs/105/renamed-05.vcf.gz > outputs/123/vcfs/lca15-region.vcf
bcftools view -r NC_066847.1:13250000-13450000 outputs/105/renamed-05.vcf.gz > outputs/123/vcfs/lca15-region1.vcf
bcftools view -r NC_066847.1:17500000-18250000 outputs/105/renamed-05.vcf.gz > outputs/123/vcfs/lca15-region2.vcf
bcftools view -r NC_066847.1:27700000-27800000 outputs/105/renamed-05.vcf.gz > outputs/123/vcfs/lca15-region3.vcf
bcftools view -r NC_066847.1 outputs/105/renamed-05.vcf.gz > outputs/123/vcfs/lca15.vcf

for f in *.vcf; do plink --vcf $f --r2 inter-chr --ld-window-r2 0.2 --out `basename $f vcf`ldf --allow-extra-chr --double-id; done;

```


```{r}
ld<-read.delim("outputs/123/vcfs/lca15-region.ldf.ld",sep="",stringsAsFactors=FALSE) %>% as_tibble() %>% dplyr::arrange(R2) %>%  filter(R2 >0.3)

ldregion<-ggplot(ld) +
  geom_point(aes(x=BP_A, y=BP_B, color=R2), alpha=0.5) +
  scale_color_gradient(low="khaki1", high="red") + 
  ggtitle(paste0(chrom))+
  theme_bw() +
  theme(panel.grid = element_blank()) +
  xlab("SNP Position A") +
  ylab("SNP Position B")
ldregion
ggsave("outputs/123/lca15-high-ld-regions.jpeg")
```


The variants are highly correlated across these blocks

```{r}
ld<-read.delim("outputs/123/vcfs/lca15-region1.ldf.ld",sep="",stringsAsFactors=FALSE) %>% as_tibble() %>% dplyr::arrange(R2) %>%  filter(R2 >0.3)

ld1<-ggplot(ld) +
  geom_point(aes(x=BP_A, y=BP_B, color=R2), alpha=0.5) +
  scale_color_gradient(low="khaki1", high="red") + 
  ggtitle(paste0(chrom))+
  theme_bw() +
  theme(panel.grid = element_blank()) +
  xlab("SNP Position A") +
  ylab("SNP Position B")
ld1
ggsave("outputs/123/lca15-high-ld-region-1.jpeg")
```

```{r}
ld<-read.delim("outputs/123/vcfs/lca15-region2.ldf.ld",sep="",stringsAsFactors=FALSE) %>% as_tibble() %>% dplyr::arrange(R2) %>%  filter(R2 >0.3)

ld2<-ggplot(ld) +
  geom_point(aes(x=BP_A, y=BP_B, color=R2), alpha=0.5) +
  scale_color_gradient(low="khaki1", high="red") + 
  ggtitle(paste0(chrom))+
  theme_bw() +
  theme(panel.grid = element_blank()) +
  xlab("SNP Position A") +
  ylab("SNP Position B")
ld2
ggsave("outputs/123/lca15-high-ld-region-2.jpeg")
```

Let's look at a pca


```{r}
vcf<-read.vcfR(file="outputs/123/vcfs/lca15-region1.vcf")

genind<-vcfR2genind(vcf)
genind@pop<-as.factor(mq$NewPhenotype)

gen<-genind
X <- tab(gen, NA.method="mean")

pca1 <- dudi.pca(X,scannf=FALSE,scale=FALSE,nf=3)

df<-pca1$li
df<-df %>% bind_cols(mq)
eig<-pca1$eig/sum(pca1$eig)*100


pc12<-ggplot(df) + 
  geom_point(aes(x=Axis1, y=Axis2, fill=NewPhenotype),pch=21, alpha=0.75, cex=2) +
  theme_bw() +
  xlab(paste0("PC1 ",round(eig[1],2), "%")) +
  ylab(paste0("PC2 ",round(eig[2],2), "%")) +
  theme(panel.grid = element_blank()) +
  guides(fill = guide_legend(override.aes = list(pch = 22))) +
  theme(legend.position = "top") +
  scale_fill_manual(values=c("red","blue","grey"))
pc12

```

loadings

```{r}
loads<-pca1$c1 %>% as_tibble()
loads$locus<-rownames(pca1$c1)
loads %>% arrange(-CS1) %>% relocate(locus) %>% filter(CS1 > 0.07) %>% write_csv("outputs/123/loadings.csv")
```

```{r}
snps<-import.snpR.data(genotypes = "outputs/123/vcfs/lca15-region1.vcf", 
      sample.meta = mq %>% dplyr::select(specimen_id,NewPhenotype))
```

```{r}
hets<-calc_hs(snps)
ho<-get.snpR.stats(hets, stats="hs")
h2o<-ho$sample %>% dplyr::select(specimen_id, NewPhenotype, hs)
h2o
df2<-df %>% left_join(h2o)
```
```{r}
hplot<-ggplot(df2) + 
  geom_point(aes(x=Axis1, y=Axis2, fill=hs),pch=21, alpha=0.75, cex=2) +
  theme_bw() +
  xlab(paste0("PC1 ",round(eig[1],2), "%")) +
  ylab(paste0("PC2 ",round(eig[2],2), "%")) +
  theme(panel.grid = element_blank()) +
  guides(fill = guide_legend(override.aes = list(pch = 22))) +
  theme(legend.position = "right") +
  scale_fill_viridis_c(option="viridis")
hplot
```

```{r}
df3<-df2 %>% mutate(Genotype=ifelse(Axis1 < -8, "Ancestral", ifelse(Axis1 >= -8 & Axis1 <= 8, "Heterozygote","Derived")))
df3$Genotype<-factor(df3$Genotype, levels=c("Ancestral","Heterozygote","Derived"))
df3<-df3 %>% mutate(Category=ifelse(NewPhenotype=="SFF","Migratory",ifelse(NewPhenotype=="SSS","Resident","UNK")))
df3 %>% ggplot() + geom_point(aes(x=Axis1, y=Axis2, fill=Genotype),pch=21, alpha=0.75, cex=2) 
```
```{r}
hplotb<-ggplot(df3) +
  geom_boxplot(aes(x=Genotype, y=hs, fill=Genotype)) +
  theme_bw() +
  ylab("Individual Heterozygosity Hs") +
  theme(panel.grid = element_blank())  +
  scale_fill_viridis_d(option="magma", alpha=0.7) +
  theme(legend.position = "none")
hplotb 
```


Let's look at genotype frequencies from the burdekin

```{r}
df3 %>% filter(Group=="Burdekin") %>% group_by(NewPhenotype, Genotype) %>% summarize(Count=n()) %>% ungroup() %>% group_by(NewPhenotype) 
```

F(A) in catadromous fish = (2+9)/42 = 0.26
F(A) in resident fish    = (14+12)/46 = 0.565

```{r}
pcaplot<-df3 %>% filter(Group=="Burdekin") %>% ggplot() + geom_point(aes(x=Axis1, y=Axis2, fill=Genotype),pch=21, alpha=0.75, cex=2) +
  facet_grid(Category~.) +
  theme_bw() +
  xlab(paste0("PC1 ",round(eig[1],2), "%")) +
  ylab(paste0("PC2 ",round(eig[2],2), "%")) +
  theme(panel.grid = element_blank()) +
  guides(fill = guide_legend(override.aes = list(pch = 22))) +
  theme(legend.position = "bottom") +
  scale_fill_viridis_d(option="magma")

pcaplot
```



```{r}
df3<-df2 %>% mutate(Genotype=ifelse(Axis1 < -8, "Ancestral", ifelse(Axis1 >= -8 & Axis1 <= 8, "Heterozygote","Derived")))
df3$Genotype<-factor(df3$Genotype, levels=c("Ancestral","Heterozygote","Derived"))

df3 %>% ggplot() + geom_point(aes(x=Axis1, y=Axis2, fill=Genotype),pch=21, alpha=0.75, cex=2) 


```
```{r}
hplotb<-ggplot(df3) +
  geom_boxplot(aes(x=Genotype, y=hs, fill=Genotype)) +
  theme_bw() +
  ylab("Individual Heterozygosity Hs") +
  theme(panel.grid = element_blank())  +
  scale_fill_viridis_d(option="magma", alpha=0.7) +
  theme(legend.position = "none")
hplotb 

df3 %>% group_by(Genotype) %>% summarize(Count=n())
```
Calculate the fst difference

```{r}

df4<-df3 %>% as_tibble()
mq2<-read_csv("meta/mq.csv")
mq2$Group<-gsub(" ","",mq2$Group)
ci<- mq2 %>% dplyr::select(specimen_id,NewPhenotype,Group) %>% left_join(df3 %>% select(specimen_id,NewPhenotype,Genotype,hs, Axis1, Axis2))
#"outputs/123/vcfs/lca15-region1.vcf", 
snps2<-import.snpR.data(genotypes = "outputs/123/vcfs/lca15.vcf",
      sample.meta = ci)

snps3<-snps2[,snps2@sample.meta$Genotype!="Heterozygote"]
```

This is all messed up
```{r}
plot_clusters(snps3, facet = "Genotype", plot_type = "pca", interpolation_method = "af")
```

```{r}
df3 %>% select(specimen_id, Genotype) %>% write_csv("meta/mq-genos.csv")
```

```{r}
fst<-calc_global_fst(snps2, "Genotype")
fsts<-get.snpR.stats(fst, facets="Genotype",stats = "fst")
fdf<-fsts$pairwise %>% as_tibble()
```

```{r}
fdf %>% ggplot() +
  geom_point(aes(x=position, y=fst, fill=fst), pch=21) +
  scale_fill_viridis_c()
```

Let's look at genotype frequencies from the burdekin

```{r}
df3 %>% filter(Group=="Burdekin") %>% group_by(NewPhenotype, Genotype) %>% summarize(Count=n()) %>% ungroup() %>% group_by(NewPhenotype) 
```

F(A) in catadromous fish = (2+9)/44 = 0.25
F(A) in resident fish    = (14+12)/44 = 0.59

mutate(Error=sqrt((Frequency*(1-Frequency))/(2*N )))
```{r}
pheno<-c("Migratory","Resident")
Frequency<-c(((2*1)+9)/(2+18+24),((2*7)+12)/(14+24+6))
sizes<-c(((1+9+12)*2),((7+12+3)*2))


t1<-cbind(pheno, Frequency, sizes) %>% as_tibble() 
t1$Frequency<-as.numeric(t1$Frequency)
t1$sizes<-as.numeric(t1$sizes)
t1 <- t1 %>% mutate(N = sizes/2) %>% mutate(Error=sqrt((Frequency*(1-Frequency))/(2*N )))

fplot<-t1 %>% ggplot() +
   geom_segment(aes(x=pheno, xend=pheno, y=Frequency-Error, yend=Frequency+Error)) +
    geom_point(aes(x=pheno, y=Frequency, fill=pheno), pch=21, cex=2) +

  ylim(0,1) +
  theme_bw() +
  ylab("Frequency of Ancestral Form") +
  xlab("Phenotype") +
  scale_fill_manual(values=c("red","blue")) +
  theme(legend.position = "none")

fplot
```
```{r}
everyone<-df3  %>% group_by(Group, NewPhenotype, Genotype) %>% summarize(Count=n()) %>% ungroup() %>% pivot_wider(id_cols = c("Group", "NewPhenotype"), names_from = Genotype, values_from = Count)

everyone[is.na(everyone)]<-0
everyone %>%  group_by(Group, NewPhenotype) %>% mutate(AncestralCount=(sum((Ancestral*2)+Heterozygote))) %>% mutate(Total=2*(Ancestral+Heterozygote+Derived)) %>% mutate(AncestralFeq=AncestralCount/Total) 

```
```{r}
prop.test(x=c(11, 26), n=c(44,44), alternative = "two.sided")

```

```{r}
ggarrange(chromld + ggtitle("A") + theme(plot.title=element_text(face="bold")), 
          ldregion + ggtitle("B") + theme(plot.title=element_text(face="bold")),
          ld1 + ggtitle("C") + theme(plot.title=element_text(face="bold")), 
          ld2 + ggtitle("D") + theme(plot.title=element_text(face="bold")))


ggsave("outputs/123/lca-15-ld-plots.jpeg", width=10, height=7)
```

PCAs and heterozygosity plots

```{r}
ggarrange(hplot + ggtitle("A") + theme(plot.title=element_text(face="bold")) + labs(fill="Hs"),
          hplotb + ggtitle("B") + theme(plot.title=element_text(face="bold")) + ylab("\nIndividual Heterozygosity (Hs)"),
          pcaplot + ggtitle("C") + theme(plot.title=element_text(face="bold")), 
          fplot + ggtitle("D") + ylab("\nFrequency of Ancestral Form"),
          heights = c(1,1))

ggsave("outputs/123/lca-15-pca-plots.jpeg", width=9, height=8)

```


## Region 2

Let's look at a pca


```{r}
vcf<-read.vcfR(file="outputs/123/vcfs/lca15-region2.vcf")

genind<-vcfR2genind(vcf)
genind@pop<-as.factor(mq$NewPhenotype)

gen<-genind
X <- tab(gen, NA.method="mean")

pca1 <- dudi.pca(X,scannf=FALSE,scale=FALSE,nf=3)

df<-pca1$li
df<-df %>% bind_cols(mq)
eig<-pca1$eig/sum(pca1$eig)*100


pc12<-ggplot(df) + 
  geom_point(aes(x=Axis1, y=Axis2, fill=NewPhenotype),pch=21, alpha=0.75, cex=2) +
  theme_bw() +
  xlab(paste0("PC1 ",round(eig[1],2), "%")) +
  ylab(paste0("PC2 ",round(eig[2],2), "%")) +
  theme(panel.grid = element_blank()) +
  guides(fill = guide_legend(override.aes = list(pch = 22))) +
  theme(legend.position = "top") +
  scale_fill_manual(values=c("red","blue","grey"))
pc12

```

```{r}
snps<-import.snpR.data(genotypes = "outputs/123/vcfs/lca15-region2.vcf", 
      sample.meta = mq %>% dplyr::select(specimen_id,NewPhenotype))
```

```{r}
hets<-calc_hs(snps)
ho<-get.snpR.stats(hets, stats="hs")
h2o<-ho$sample %>% dplyr::select(specimen_id, NewPhenotype, hs)
h2o
df2<-df %>% left_join(h2o)
```
```{r}
hplot<-ggplot(df2) + 
  geom_point(aes(x=Axis1, y=Axis2, fill=hs),pch=21, alpha=0.75, cex=2) +
  theme_bw() +
  xlab(paste0("PC1 ",round(eig[1],2), "%")) +
  ylab(paste0("PC2 ",round(eig[2],2), "%")) +
  theme(panel.grid = element_blank()) +
  guides(fill = guide_legend(override.aes = list(pch = 22))) +
  theme(legend.position = "right") +
  scale_fill_viridis_c(option="viridis")
hplot
```

```{r}
df3<-df2 %>% mutate(Genotype=ifelse(Axis1 < -8, "Ancestral", ifelse(Axis1 >= -8 & Axis1 <= 8, "Heterozygote","Derived")))
df3$Genotype<-factor(df3$Genotype, levels=c("Ancestral","Heterozygote","Derived"))

df3 %>% ggplot() + geom_point(aes(x=Axis1, y=Axis2, fill=Genotype),pch=21, alpha=0.75, cex=2) 


```
```{r}
hplotb<-ggplot(df3) +
  geom_boxplot(aes(x=Genotype, y=hs, fill=Genotype)) +
  theme_bw() +
  ylab("Individual Heterozygosity Hs") +
  theme(panel.grid = element_blank())  +
  scale_fill_viridis_d(option="magma", alpha=0.7) +
  theme(legend.position = "none")
hplotb 
```


Let's look at genotype frequencies from the burdekin

```{r}
df3 %>% filter(Group=="Burdekin") %>% group_by(NewPhenotype, Genotype) %>% summarize(Count=n()) %>% ungroup() %>% group_by(NewPhenotype) 
```

F(A) in catadromous fish = (2+9)/42 = 0.26
F(A) in resident fish    = (14+12)/46 = 0.565

```{r}
pcaplot<-df3 %>% filter(Group=="Burdekin") %>% ggplot() + geom_point(aes(x=Axis1, y=Axis2, fill=Genotype),pch=21, alpha=0.75, cex=2) +
  facet_grid(NewPhenotype~.) +
  theme_bw() +
  xlab(paste0("PC1 ",round(eig[1],2), "%")) +
  ylab(paste0("PC2 ",round(eig[2],2), "%")) +
  theme(panel.grid = element_blank()) +
  guides(fill = guide_legend(override.aes = list(pch = 22))) +
  theme(legend.position = "bottom") +
  scale_fill_viridis_d(option="magma")

pcaplot
```


## Region 3

Let's look at a pca


```{r}
vcf<-read.vcfR(file="outputs/123/vcfs/lca15-region3.vcf")

genind<-vcfR2genind(vcf)
genind@pop<-as.factor(mq$NewPhenotype)

gen<-genind
X <- tab(gen, NA.method="mean")

pca1 <- dudi.pca(X,scannf=FALSE,scale=FALSE,nf=3)

df<-pca1$li
df<-df %>% bind_cols(mq)
eig<-pca1$eig/sum(pca1$eig)*100


pc12<-ggplot(df) + 
  geom_point(aes(x=Axis1, y=Axis2, fill=NewPhenotype),pch=21, alpha=0.75, cex=2) +
  theme_bw() +
  xlab(paste0("PC1 ",round(eig[1],2), "%")) +
  ylab(paste0("PC2 ",round(eig[2],2), "%")) +
  theme(panel.grid = element_blank()) +
  guides(fill = guide_legend(override.aes = list(pch = 22))) +
  theme(legend.position = "top") +
  scale_fill_manual(values=c("red","blue","grey"))
pc12

```

```{r}
snps<-import.snpR.data(genotypes = "outputs/123/vcfs/lca15-region3.vcf", 
      sample.meta = mq %>% dplyr::select(specimen_id,NewPhenotype))
```

```{r}
hets<-calc_hs(snps)
ho<-get.snpR.stats(hets, stats="hs")
h2o<-ho$sample %>% dplyr::select(specimen_id, NewPhenotype, hs)
h2o
df2<-df %>% left_join(h2o)
```
```{r}
hplot<-ggplot(df2) + 
  geom_point(aes(x=Axis1, y=Axis2, fill=hs),pch=21, alpha=0.75, cex=2) +
  theme_bw() +
  xlab(paste0("PC1 ",round(eig[1],2), "%")) +
  ylab(paste0("PC2 ",round(eig[2],2), "%")) +
  theme(panel.grid = element_blank()) +
  guides(fill = guide_legend(override.aes = list(pch = 22))) +
  theme(legend.position = "right") +
  scale_fill_viridis_c(option="viridis")
hplot
```

High Fst loci between alternative homokaryotypes.

```{r}

```
