---
title: "124-local-pca-redo"
output: html_document
date: "2026-07-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Local PCA
```{r}
mq<-read_csv("meta/mq.csv")
meta<-mq
mq %>% filter(Group == "Burdekin") %>% dplyr::select(specimen_id) %>% write_tsv("outputs/124/burdekin.names", col_names = FALSE)
```


gunzip -c  outputs/105/renamed-05.vcf.gz  |  grep "#" > outputs/107/header.txt 

gunzip -c outputs/105/renamed-05.vcf.gz | grep -v "#" | awk 'BEGIN{i=0}{i++;if (i%50==0) print}' > outputs/107/t50.vcf
#22707 snps

gunzip -c outputs/105/renamed-05.vcf.gz | grep -v "#" | awk 'BEGIN{i=0}{i++;if (i%20==0) print}' > outputs/107/t20.vcf
#56768

gunzip -c outputs/105/renamed-05.vcf.gz | grep -v "#" | awk 'BEGIN{i=0}{i++;if (i%10==0) print}' > outputs/107/t10.vcf
#

cat outputs/107/header.txt  outputs/107/t50.vcf > outputs/107/thinned50.vcf; bgzip outputs/107/thinned50.vcf; tabix outputs/107/thinned50.vcf.gz;
cat outputs/107/header.txt  outputs/107/t20.vcf > outputs/107/thinned20.vcf; bgzip outputs/107/thinned20.vcf; tabix outputs/107/thinned20.vcf.gz;
cat outputs/107/header.txt  outputs/107/t10.vcf > outputs/107/thinned10.vcf; bgzip outputs/107/thinned10.vcf; tabix outputs/107/thinned10.vcf.gz;

```{sh, eval=FALSE}
bcftools view -S outputs/124/burdekin.names outputs/105/renamed-05.vcf.gz | bcftools +fill-tags | bcftools view -i 'INFO/MAF > 0.10' > outputs/124/burd.filt.vcf 
bgzip outputs/124/burd.filt.vcf
tabix outputs/124/burd.filt.vcf.gz

gunzip -c  outputs/124/burd.filt.vcf.gz  |  grep "#" > outputs/124/header.txt 
gunzip -c outputs/124/burd.filt.vcf.gz | grep -v "#" | awk 'BEGIN{i=0}{i++;if (i%10==0) print}' > outputs/124/t10.vcf

cat outputs/124/header.txt  outputs/124/t10.vcf > outputs/124/thinned10.vcf; bgzip outputs/124/thinned10.vcf; tabix outputs/124/thinned10.vcf.gz;

cat meta/lates-lgs.txt  | while read line; do bcftools view -Ob -r $line outputs/124/thinned10.vcf.gz > outputs/124/bcf/$line.bcf; done;

for f in outputs/124/bcf/*.bcf; do bcftools index $f; done;
```

```{r}
samples<-meta %>% dplyr::select(sample_id) %>% rename(ID=sample_id)

population<-meta %>% dplyr::select(Group) %>% rename(population=Group)

table<-cbind(samples, population)
write.table(table, "outputs/124/bcf/sample_info.tsv", quote = TRUE, row.names = FALSE, sep="\t")

```


thinned 10 82064 variants
```{sh, eval=FALSE}
./run_lostruct.R -i /Users/mac/github/barra-lh/outputs/124/bcf -t snp -s 20 -m 4 -I /Users/mac/github/barra-lh/outputs/124/bcf/sample_info.tsv -j 124
cp lostruct_results/type_snp_size_20_weights_none_jobid_124/mds_coords.csv ~/github/barra-lh/outputs/124/mds_coords-10.csv

./run_lostruct.R -i /Users/mac/github/barra-lh/outputs/124/bcf -t snp -s 10 -m 4 -I /Users/mac/github/barra-lh/outputs/124/bcf/sample_info.tsv -j 124
cp lostruct_results/type_snp_size_10_weights_none_jobid_124/mds_coords.csv ~/github/barra-lh/outputs/124/mds_coords-10-10.csv
```

```{r}
mds<-read_csv("outputs/124/mds_coords-10.csv") # 20 snp windows, thinned 10

#make tidy
tidymds<-mds %>% gather(MDS, Value, 3:6)
MDS1<-filter(tidymds, MDS=="MDS1") %>% rename(MDS1=MDS) %>% rename(Value1=Value)
MDS2<-filter(tidymds, MDS=="MDS2") %>% rename(MDS2=MDS) %>% rename(Value2=Value)
MDS3<-filter(tidymds, MDS=="MDS3") %>% rename(MDS3=MDS) %>% rename(Value3=Value)
MDS4<-filter(tidymds, MDS=="MDS4") %>% rename(MDS4=MDS) %>% rename(Value4=Value)
```

```{r}
ggplot(mds)+
  geom_point(aes(x=MDS1, y=MDS2, fill=chrom), pch=21, alpha=0.75) +
  scale_fill_viridis_d(option="magma") +
  theme_bw() +
  theme(legend.position = 'none')


ggplot(mds)+
  geom_point(aes(x=MDS1, y=MDS3, fill=chrom), pch=21, alpha=0.75) +
  scale_fill_viridis_d(option="magma") +
  theme_bw() +
  theme(legend.position = 'none')


ggplot(mds)+
  geom_point(aes(x=MDS1, y=MDS4, fill=chrom), pch=21, alpha=0.75) +
  scale_fill_viridis_d(option="magma") +
  theme_bw() +
  theme(legend.position = 'none')
```

```{r}
p1<-MDS1 %>% mutate(Index=1:n())

out <- boxplot.stats(p1$Value1)$out
out_ind <- which(p1$Value1 %in% c(out))
length(out_ind)
outliers<-p1[out_ind,]
outliers %>% group_by(chrom) %>% summarize(Count=n()) %>% arrange(-Count)
```

```{r}
chroms<-p1 %>% group_by(chrom) %>% mutate(Start=min(Index), Stop=max(Index)) %>% select(chrom,Start,Stop) %>% unique() %>% 
  mutate(Position=round((Start+Stop)/2))

#Make enough colors
nb.cols <- nrow(chroms)
mycolors <- colorRampPalette(brewer.pal(8, "Set1"))(nb.cols)

ggplot(p1) +
  geom_rect(data=chroms, aes(xmin=Start, xmax=Stop, ymin=min(p1$Value1), ymax=max(p1$Value1)), fill=mycolors, alpha=0.25) +
  geom_point(data=p1, aes(x=Index, y=Value1, color=chrom), alpha=0.75, cex=0.5) +
  geom_point(data=outliers, aes(x=Index, y=Value1), color="black", cex=0.5) +
  theme_bw() +
  scale_x_continuous(breaks=chroms$Position, labels=chroms$chrom) +
  theme(axis.text.x = element_text(angle=45, hjust = 1, face="bold", size=6)) +
  theme(legend.title = element_blank()) +
  theme(panel.grid = element_blank()) +
  theme(legend.position = "none") +
  scale_color_manual(values=mycolors) +
  ylab("MDS1") +
  xlab("Chromosome")
```